---
title: "Replication Challenge"
author: "Lisa Philine Krämer"
date: "2025-11-20"
output: html_document
---
# Setup ----
```{r}
library(tidyverse)
library(haven)
library(ggplot2)
library(dplyr)
library(plotly)
library(tidyr)
library(xtable)
library(knitr)
library(kableExtra)
library(tidybayes)
library(brms)
library(bayesplot)
library(posterior)
library(dagitty)
library(ggdag)
library(cmdstanr)
library(brms)
library(loo)
library(brms)
library(bayesplot)
```

```{r}
data <- read_dta("original&computed_DB_forrep.dta")
```

# Data Preparation and Exploration ----
```{r}
# sort the columns to be alphabetically ordered
data <- data[, order(names(data))]

# change dates to Date type 
data$bdate <- as.Date(data$bdate, format = "%Y-%m-%d")
data$joineddate <- as.Date(data$joineddate, format = "%Y-%m-%d")
```

```{r}
summary(data)
```

```{r}
data %>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
  facet_wrap(~ key, scales = "free") +   # In separate panels
  geom_histogram() 
```



```{r}
# Count of NAs per column
colSums(is.na(data))

# Percentage of missing values
colSums(is.na(data)) / nrow(data) * 100

```

```{r}
#plot every variable's histogram, since all data is either numeric or date

all_vars <- names(data)

for (v in all_vars) {
  print(
    ggplot(data, aes_string(x = v)) +
      geom_histogram(bins = 30) +
      ggtitle(paste("Histogram of", v))
  )
}


```

```{r}
# heatmap


# make a numeric-only version of your data:
data_num <- data %>%
  mutate(across(
    where(~ inherits(.x, "Date")),  # <- predicate: is it a Date?
    as.numeric
  ))

# calculate correlations
cor_mat <- cor(data_num, use = "pairwise.complete.obs", method = "pearson")

library(plotly)

# custom red–white–blue colorscale: blue (negative) → white (0) → red (positive)
rwb_scale <- list(
  list(0.0, "blue"),
  list(0.5, "white"),
  list(1.0, "red")
)

plot_ly(
  x = colnames(cor_mat),
  y = rownames(cor_mat),
  z = cor_mat,
  type = "heatmap",
  zmin = -1,
  zmax = 1,
  colorscale = rwb_scale,
  hovertemplate = paste(
    "Var1: %{x}<br>",
    "Var2: %{y}<br>",
    "Correlation: %{z:.2f}<extra></extra>"
  )
) %>%
  layout(
    title = list(
      text = "Pearson Correlation Heatmap",
      x = 0.5,
      xanchor = "center"
    ),
    xaxis = list(
      title = "",
      tickangle = -45,
      tickfont = list(size = 9)
    ),
    yaxis = list(
      title = "",
      tickfont = list(size = 9,
                      automargin = TRUE)
    ),
    margin = list(l = 80, r = 20, b = 120, t = 50),
    coloraxis = list(
      colorbar = list(
        title = "r",
        ticksuffix = "",
        len = 0.8
      )
    )
  )


```


  1##Prepare Data Sets
```{r}
data_base <- data %>%
  filter(
    dfullsection == 1,          # full sections only
    position %in% c(3, 4),      # midfielders + strikers
    !(section %in% c(10, 20)) 
  )
data_base <- data_base  %>% mutate(player = factor(player),
                                   gamenu = factor(gamenu),
                                   section = factor(section))
```

## DAGs
```{r}
dag1 <- dagitty("dag{
                S -> G
                NoS -> G
                NoS -> S 
                E -> S
                E -> NoS}")

coordinates(dag1) <- list( 
                          x = c(S = 0, NoS = 1, G = 2, E = 0) , 
                          y = c(S = 0, NoS = 1, G = 0, E = 1))

ggdag( dag1 ) + 
  ggtitle( "(1) Effort influences Speed & Number of Sprints;\n(2) Speed influences Goal Probability;\n(3) Number of Sprints influences Speed & Goal Probability" ) +
  theme_dag()



dag2 <- dagitty("dag{
                D -> W
                OD -> W
                D -> R
                OD -> R
                D -> T
                OD -> T
                R -> W
                T -> W
                E -> D
                OE -> OD}")

coordinates(dag2) <- list( 
                          x = c(D = 0, OD = 2, W = 1, R = 1, "T" = 1, E = -1, OE = 3) , 
                          y = c(D = 1, OD = 1, W = 1, R = 2, "T" = 0, E = 1, OE = 1))

ggdag( dag2 ) + 
  ggtitle( "(1) Teams Effort influences Team Distance;\n(2) Opponents Effort influences Opponents Distance;\n(3) Team Distance influences Team Winning Probability;\n(4) Opponent Distance influences Team Winning Probability;\n(5) Team Distance and Opponents Distance influence the Distance Ratio and the Total Distance" ) +
  theme_dag()
```

```{r}
library(fixest)

reg1 <- feols(dplagoal ~ plaavspeed, data = data_base)
reg2 <- feols(dplagoal ~ playersprints, data = data_base)
reg3 <- feols(dplagoal ~ plaavspeed + playersprints, data = data_base)
reg4 <- feols(plaavspeed ~ playersprints, data = data_base)

etable(reg1, reg2, reg3, reg4)

```

```{r}
data %>%
  ggplot(aes(x=playersprints, y=plaavspeed)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "steelblue") +
  labs(x = "Number of Sprints",
       y = "Speed (km/h)") +
  theme_classic()


data %>%
  ggplot(aes(x=dplagoal)) +
  geom_histogram(fill = "steelblue", binwidth = 0.1) +
  labs(x = "Goal Probability",
        y = "Count") +
  theme_classic()


data %>%
  ggplot(aes(x=dteamwin)) +
  geom_histogram(fill = "steelblue", binwidth = 0.1) +
  labs(x = "Team Winning Probability",
        y = "Count") +
  theme_classic()


data %>%
  ggplot(aes(x=p_odds_team, y = dteamwin)) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "steelblue") +
  labs(x = "BO Team Win",
       y = "Team Winning Probability") +
  theme_classic()
```


## Summary Statistics

Define missing variable

```{r}
team_game_redcards <- data_base %>%
  group_by(team, gamenu) %>%
  summarise(
    redcard_teamgame = as.integer(any(dplaredc == 1)),
    .groups = "drop"
  )

```

```{r}
mean(data$dfullsection, na.rm = TRUE)
```

```{r}
data_PA <- data %>%
  filter(
    dfullsection == 1,               # full sections only
    !(section %in% c(10, 20))) %>%   # no overtime sections
  mutate(
    SMGP = ifelse(position %in% c(3,4),          # SMGP = Strikers and midfielders goal probability
                  dplagoal,                      # already 0/1 if player scored
                  NA))

```


```{r}
### --------------------------------------------------------------
### PANEL A: Section level
### --------------------------------------------------------------

panelA_vars <- list(
  full_sections                = "dfullsection",
  overtime_sections            = "(data$section %in% c(10, 20)),2)",
  player_distance              = "playerdist",
  player_sprints               = "playersprints",
  goal_probability             = "SMGP",
  team_distance                = "teamdist"
)

panelA <- data_PA %>% 
  summarise(
    FullSections_mean     = mean(data$dfullsection, na.rm = TRUE),
    Overtime_mean         = round(mean(data$section %in% c(10, 20)),2),
    PlayerDistance_mean   = mean(playerdist, na.rm = TRUE),
    PlayerDistance_sd     = sd(playerdist, na.rm = TRUE),
    PlayerSprints_mean    = mean(playersprints, na.rm = TRUE),
    PlayerSprints_sd      = sd(playersprints, na.rm = TRUE),
    GoalProb_mean         = mean(SMGP, na.rm = TRUE),
    GoalProb_sd           = sd(SMGP, na.rm = TRUE),
    TeamDistance_mean     = data %>% filter(!(section %in% c(10, 20))) %>% summarise(mean(teamdist, na.rm = TRUE)) %>% pull(),
    TeamDistance_sd       = data %>% filter(!(section %in% c(10, 20))) %>% summarise(sd(teamdist, na.rm = TRUE)) %>% pull(),
    Observations_PS       = nrow(data),
    Observations_TS       = data %>% distinct(team, gamenu) %>% nrow()*20   # unique combinations of team and games      

  )

options(scipen = 999)
panelA <- panelA %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )

```

```{r}

dataB <- data %>%
  group_by(gamenu, player) %>%
  summarise(
    player_sectionsplayed = first(player_sectionsplayed),  
    player_avg_speed      = mean(plaavspeed_game, na.rm = TRUE),
    .groups = "drop"
  )

teamdistance <- data %>%
  group_by(gamenu, team) %>%
  summarise(
    teamdist_game = mean(teamdist_game, na.rm = TRUE)
  )

crowd_game <- data %>%
  group_by(gamenu, team, dhomegame) %>%  
  summarise(
    home_crowd = first(team_crowd),
    opp_crowd  = first(opp_crowd),
    .groups = "drop"
  )

redcard <- data %>%
  group_by(gamenu) %>%
  summarise(
    any_red = max(dplaredc, na.rm = TRUE),  
    .groups = "drop"
  )

```



```{r}
### --------------------------------------------------------------
### PANEL B: Game level
### --------------------------------------------------------------

panelB <- data %>% 
  summarise(
    SectionsPlayed_mean   = mean(dataB$player_sectionsplayed, na.rm = TRUE),
    SectionsPlayed_sd     = sd(dataB$player_sectionsplayed, na.rm = TRUE),
    AvgSpeed_mean         = mean(dataB$player_avg_speed, na.rm = TRUE),
    AvgSpeed_sd           = sd(dataB$player_avg_speed, na.rm = TRUE),
    WinningProb_mean      = mean(data$dteamwin),
    WinningProb_sd        = sd(data$dteamwin),
    TeamDistanceGame_mean = mean(teamdistance$teamdist_game, na.rm = TRUE),
    TeamDistanceGame_sd   = sd(teamdistance$teamdist_game, na.rm = TRUE),
    HomeCrowd_mean        = mean(crowd_game$home_crowd[crowd_game$dhomegame == 1], na.rm = TRUE),
    HomeCrowd_sd          = sd(crowd_game$home_crowd[crowd_game$dhomegame == 1], na.rm = TRUE),
    AwayCrowd_mean        = mean(crowd_game$opp_crowd[crowd_game$dhomegame == 1], na.rm = TRUE),
    AwayCrowd_sd          = sd(crowd_game$opp_crowd[crowd_game$dhomegame == 1], na.rm = TRUE),
    RedCardProb_mean      = mean(redcard$any_red, na.rm = TRUE),
    RedCardProb_sd        = sd(redcard$any_red, na.rm = TRUE),
    BO_mean               = mean(data$p_odds_team, na.rm = T),
    BO_sd                 = sd(data$p_odds_team, na.rm = T),
    Obs_player_game       = data %>% distinct(gamenu, player) %>% nrow(),
    Obs_team_game         = n_distinct(team, gamenu),
    
    Rounds                = n_distinct(played),
    Games                 = n_distinct(gamenu),
    Teams                 = n_distinct(team),
    Players               = n_distinct(player),
    Strikers              = n_distinct(data$player[data$position == 4]),
    Midfielders           = n_distinct(data$player[data$position == 3]),
    Defenders             = n_distinct(data$player[data$position == 2])
  )

panelB <- panelB %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )
```

## Latex code for summary statistics
```{r}
### PANEL A Latex code
panelA_latex <- kable(panelA, 
                      format = "latex", 
                      booktabs = TRUE, 
                      caption = "Panel A: Section LeveL") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))

# Print Latex code
cat(panelA_latex)

### PANEL B Latex code
panelB_latex <- kable(panelB, 
                      format = "latex", 
                      booktabs = TRUE, 
                      caption = "Panel B: Game Level") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))

# Print Latex code
cat(panelB_latex)
```



# Data Preparation ----
```{r}
data_col1 <- data_base %>%
  drop_na(dplagoal, plaavspeed) %>%
  mutate(
    plaavspeed_c = plaavspeed - mean(plaavspeed, na.rm = TRUE)
  )
```

```{r}
data_col2 <- data_base %>%
  drop_na(
    dplagoal, plaavspeed,
    dhomegame, rank_gap,
    player_cumsections, player_cumsections_2
  ) %>%
  mutate(
    plaavspeed_c           = plaavspeed - mean(plaavspeed, na.rm = TRUE),
    rank_gap_c             = rank_gap - mean(rank_gap, na.rm = TRUE),
    player_cumsections_c   = player_cumsections - mean(player_cumsections, na.rm = TRUE),
    player_cumsections_2_c = player_cumsections_2 - mean(player_cumsections_2, na.rm = TRUE)
  )

```

```{r}
data_col356 <- data_base %>%
  drop_na(
    dplagoal, plaavspeed, playersprints,
    dhomegame, rank_gap,
    player_cumsections, player_cumsections_2,
    scoregap_pre, dteamgoal_pre, dplagoal_pre,
    dteamred_pre, dteaminj_sec_pre,
    team_avespeed_pre
  ) %>%
  mutate(
    plaavspeed_c           = plaavspeed - mean(plaavspeed, na.rm = TRUE),
    playersprints_c        = playersprints - mean(playersprints, na.rm = TRUE),
    rank_gap_c             = rank_gap - mean(rank_gap, na.rm = TRUE),
    player_cumsections_c   = player_cumsections - mean(player_cumsections, na.rm = TRUE),
    player_cumsections_2_c = player_cumsections_2 - mean(player_cumsections_2, na.rm = TRUE),
    scoregap_pre_c         = scoregap_pre - mean(scoregap_pre, na.rm = TRUE),
    team_avespeed_pre_c    = team_avespeed_pre - mean(team_avespeed_pre, na.rm = TRUE)
  )
```

```{r}
data_col4 <- data_base %>%
  drop_na(
    dplagoal, plaavspeed,
    dhomegame, rank_gap,
    player_cumsections, player_cumsections_2,
    scoregap_pre, dteamgoal_pre, dplagoal_pre,
    dteamred_pre, dteaminj_sec_pre,
    team_avespeed_pre,
    team_crowd, opp_crowd,
    p_odds_team, p_odds_draw
  ) %>%
  mutate(
    plaavspeed_c           = plaavspeed - mean(plaavspeed, na.rm = TRUE),
    rank_gap_c             = rank_gap - mean(rank_gap, na.rm = TRUE),
    player_cumsections_c   = player_cumsections - mean(player_cumsections, na.rm = TRUE),
    player_cumsections_2_c = player_cumsections_2 - mean(player_cumsections_2, na.rm = TRUE),
    scoregap_pre_c         = scoregap_pre - mean(scoregap_pre, na.rm = TRUE),
    team_avespeed_pre_c    = team_avespeed_pre - mean(team_avespeed_pre, na.rm = TRUE),
    team_crowd_c           = team_crowd - mean(team_crowd, na.rm = TRUE),
    opp_crowd_c            = opp_crowd - mean(opp_crowd, na.rm = TRUE),
    p_odds_team_c          = p_odds_team - mean(p_odds_team, na.rm = TRUE),
    p_odds_draw_c          = p_odds_draw - mean(p_odds_draw, na.rm = TRUE)
  )

```

```{r}
# player × game data for column 7 (as in the paper)
data_col7 <- data %>%
  # same sample restrictions as Table 2
  filter(
    position %in% c(3, 4),         # only midfielders and strikers
    !(section %in% c(10, 20))      # drop overtime sections
  ) %>%
  group_by(player, gamenu) %>%
  summarise(
    # outcome: at least one goal by this player in this game
    goal_game        = as.integer(any(dplagoal == 1)),
    
    # main regressor: game-average speed of the player
    plaavspeed_game  = first(plaavspeed_game),
    
    # controls that are constant within a game for that player
    dhomegame        = first(dhomegame),
    rank_gap         = first(rank_gap),
    
    .groups = "drop"
  ) %>%
  # drop missing info (e.g. if some game-level variable is NA)
  drop_na(goal_game, plaavspeed_game, dhomegame, rank_gap) %>%
  # centering (like in the other columns)
  mutate(
    plaavspeed_game_c = plaavspeed_game - mean(plaavspeed_game, na.rm = TRUE),
    rank_gap_c        = rank_gap - mean(rank_gap, na.rm = TRUE),
    # factors for player and game fixed/random effects
    player            = factor(player),
    gamenu            = factor(gamenu)
  )

```

```{r}
#check if n matches table
nrow(data_col1)
nrow(data_col2)
nrow(data_col356) #unterschieded sich um 2, müsste aber eig gleich sein
nrow(data_col4)
nrow(data_col7) #hier fehlen 14, verstehe nicht warum
nrow(data_base)
```



# Setup Table 2 ----

```{r}
# ============================================================
# (A) HIERARCHICAL priors (for models with (1 | player) etc.)
# ============================================================

priors_base <- c(
  set_prior("normal(-4.6, 0.3)", class = "Intercept"),
  set_prior("student_t(3, 0, 1)", class = "sd", group = "player",  lb = 0),
  set_prior("student_t(3, 0, 1)", class = "sd", group = "gamenu",  lb = 0),
  set_prior("student_t(3, 0, 1)", class = "sd", group = "section", lb = 0)
)

priors_col1 <- c(
  priors_base,
  set_prior("normal(0, 0.08)", class = "b", coef = "plaavspeed_c")
)

priors_col2 <- c(
  priors_base,
  set_prior("normal(0, 0.08)",  class = "b", coef = "plaavspeed_c"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.06)",  class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c")
)

priors_col3 <- c(
  priors_base,
  set_prior("normal(0, 0.08)",  class = "b", coef = "plaavspeed_c"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.06)",  class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c"),
  set_prior("normal(0, 0.15)",  class = "b", coef = "scoregap_pre_c"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteamgoal_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dplagoal_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteamred_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteaminj_sec_pre"),
  set_prior("normal(0, 0.025)", class = "b", coef = "team_avespeed_pre_c")
)

priors_col4 <- c(
  priors_col3,
  set_prior("normal(0, 4.8e-05)", class = "b", coef = "team_crowd_c"),
  set_prior("normal(0, 4.8e-05)", class = "b", coef = "opp_crowd_c"),
  set_prior("normal(0, 1.5)",     class = "b", coef = "p_odds_team_c"),
  set_prior("normal(0, 2.5)",     class = "b", coef = "p_odds_draw_c")
)

priors_col5 <- c(
  priors_base,
  set_prior("normal(0, 0.10)",  class = "b", coef = "playersprints_c"),
  set_prior("normal(0, 0.40)",  class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.06)",  class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c"),
  set_prior("normal(0, 0.15)",  class = "b", coef = "scoregap_pre_c"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteamgoal_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dplagoal_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteamred_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteaminj_sec_pre"),
  set_prior("normal(0, 0.025)", class = "b", coef = "team_avespeed_pre_c")
)

priors_col6 <- c(
  priors_col5,
  set_prior("normal(0, 0.08)", class = "b", coef = "plaavspeed_c")
)

priors_base_col7 <- c(
  set_prior("normal(-2.2, 0.3)", class = "Intercept"),
  set_prior("student_t(3, 0, 1)", class = "sd", group = "player", lb = 0),
  set_prior("student_t(3, 0, 1)", class = "sd", group = "gamenu", lb = 0)
)

priors_col7 <- c(
  priors_base_col7,
  set_prior("normal(0, 0.05)",  class = "b", coef = "plaavspeed_game_c"),
  set_prior("normal(0, 0.5)",   class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.015)", class = "b", coef = "rank_gap_c")
)


# ============================================================
# (B) FIXED-EFFECTS priors (for models with factor(player) etc.)
# ============================================================

priors_fixef_base <- c(
  set_prior("normal(-4.6, 0.3)", class = "Intercept"),
  set_prior("normal(0, 0.4)",    class = "b")
)

priors_fixef_col1 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.08)", class = "b", coef = "plaavspeed_c")
)

priors_fixef_col2 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.08)",  class = "b", coef = "plaavspeed_c"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.06)",  class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c")
)

priors_fixef_col3 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.08)",  class = "b", coef = "plaavspeed_c"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.06)",  class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c"),
  set_prior("normal(0, 0.15)",  class = "b", coef = "scoregap_pre_c"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteamgoal_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dplagoal_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteamred_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteaminj_sec_pre"),
  set_prior("normal(0, 0.025)", class = "b", coef = "team_avespeed_pre_c")
)

priors_fixef_col4 <- c(
  priors_fixef_col3,
  set_prior("normal(0, 4.8e-05)", class = "b", coef = "team_crowd_c"),
  set_prior("normal(0, 4.8e-05)", class = "b", coef = "opp_crowd_c"),
  set_prior("normal(0, 1.5)",     class = "b", coef = "p_odds_team_c"),
  set_prior("normal(0, 2.5)",     class = "b", coef = "p_odds_draw_c")
)

priors_fixef_col5 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.10)",  class = "b", coef = "playersprints_c"),
  set_prior("normal(0, 0.40)",  class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.06)",  class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c"),
  set_prior("normal(0, 0.15)",  class = "b", coef = "scoregap_pre_c"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteamgoal_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dplagoal_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteamred_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteaminj_sec_pre"),
  set_prior("normal(0, 0.025)", class = "b", coef = "team_avespeed_pre_c")
)

priors_fixef_col6 <- c(
  priors_fixef_col5,
  set_prior("normal(0, 0.08)", class = "b", coef = "plaavspeed_c")
)

priors_fixef_base_col7 <- c(
  set_prior("normal(-2.2, 0.3)", class = "Intercept"),
  set_prior("normal(0, 0.5)",    class = "b")
)

priors_fixef_col7 <- c(
  priors_fixef_base_col7,
  set_prior("normal(0, 0.05)",  class = "b", coef = "plaavspeed_game_c"),
  set_prior("normal(0, 0.5)",   class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.015)", class = "b", coef = "rank_gap_c")
)
```

```{r}
threads_per_chain <- 2
Sys.setenv(STAN_NUM_THREADS = threads_per_chain)
```

# PP Check Fixed effects Table 2 ---- 
```{r}
m_col1_prior <- brm(
  dplagoal ~ plaavspeed_c + factor(player) + factor(gamenu) + factor(section),
  data      = data_col1,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col1,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col1 <- pp_check(m_col1_prior, ndraws = 100)
ggsave("col1_ppcheck.png", p_col1)

saveRDS(m_col1_prior, "m_col1_prior.rds")
```

```{r}
m_col2_prior <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col2,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col2,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col2 <- pp_check(m_col2_prior, ndraws = 100)
ggsave("col2_ppcheck.png", p_col2)

saveRDS(m_col2_prior, "m_col2_prior.rds")
```

```{r}
m_col3_prior <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col3,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col3 <- pp_check(m_col3_prior, ndraws = 100)
ggsave("col3_ppcheck.png", p_col3)

saveRDS(m_col3_prior, "m_col3_prior.rds")
```


```{r}
m_col4_prior <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    team_crowd_c + opp_crowd_c + p_odds_team_c + p_odds_draw_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col4,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col4,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col4 <- pp_check(m_col4_prior, ndraws = 100)
ggsave("col4_ppcheck.png", p_col4)

saveRDS(m_col4_prior, "m_col4_prior.rds")
```


# Fixed Effects Table 2 ----
```{r}
# ---- FIXED EFFECTS (Table 2), cmdstanr backend, save model + LOO/WAIC ----

# ---------- Column 1 ----------
m_col1 <- brm(
  dplagoal ~ plaavspeed_c + factor(player) + factor(gamenu) + factor(section),
  data      = data_col1,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col1,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col1 <- add_criterion(m_col1, c("loo", "waic"))
saveRDS(m_col1, "m_col1.rds")

# ---------- Column 2 ----------
m_col2 <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col2,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col2,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col2 <- add_criterion(m_col2, c("loo", "waic"))
saveRDS(m_col2, "m_col2.rds")

# ---------- Column 3 ----------
m_col3 <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col3,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col3 <- add_criterion(m_col3, c("loo", "waic"))
saveRDS(m_col3, "m_col3.rds")

# ---------- Column 4 ----------
m_col4 <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    team_crowd_c + opp_crowd_c + p_odds_team_c + p_odds_draw_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col4,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col4,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col4 <- add_criterion(m_col4, c("loo", "waic"))
saveRDS(m_col4, "m_col4.rds")

# ---------- Column 5 ----------
m_col5 <- brm(
  dplagoal ~ playersprints_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col5,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col5 <- add_criterion(m_col5, c("loo", "waic"))
saveRDS(m_col5, "m_col5.rds")

# ---------- Column 6 ----------
m_col6 <- brm(
  dplagoal ~ plaavspeed_c + playersprints_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col6,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col6 <- add_criterion(m_col6, c("loo", "waic"))
saveRDS(m_col6, "m_col6.rds")

# ---------- Column 7 ----------
m_col7 <- brm(
  goal_game ~ plaavspeed_game_c + dhomegame + rank_gap_c +
    factor(player) + factor(gamenu),
  data      = data_col7,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col7,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col7 <- add_criterion(m_col7, c("loo", "waic"))
saveRDS(m_col7, "m_col7.rds")
```

lol

# PP Check Hierarchical effects Table 2 ----
```{r}
m_col1_hier_prior <- brm(
  dplagoal ~ plaavspeed_c + (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col1,
  family    = bernoulli(link = "logit"),
  prior     = priors_col1,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col1_hier <- pp_check(m_col1_hier_prior, ndraws = 100)
ggsave("col1_ppcheck_hier.png", p_col1_hier)

saveRDS(m_col1_hier_prior, "m_col1_hier_prior.rds")

```

```{r}
m_col2_hier_prior <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col2,
  family    = bernoulli(link = "logit"),
  prior     = priors_col2,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col2_hier <- pp_check(m_col2_hier_prior, ndraws = 100)
ggsave("col2_ppcheck_hier.png", p_col2_hier)

saveRDS(m_col2_hier_prior, "m_col2_hier_prior.rds")

```

```{r}
m_col3_hier_prior <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_col3,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col3_hier <- pp_check(m_col3_hier_prior, ndraws = 100)
ggsave("col3_ppcheck_hier.png", p_col3_hier)

saveRDS(m_col3_hier_prior, "m_col3_hier_prior.rds")

```

```{r}
m_col4_hier_prior <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    team_crowd_c + opp_crowd_c + p_odds_team_c + p_odds_draw_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col4,
  family    = bernoulli(link = "logit"),
  prior     = priors_col4,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col4_hier <- pp_check(m_col4_hier_prior, ndraws = 100)
ggsave("col4_ppcheck_hier.png", p_col4_hier)

saveRDS(m_col4_hier_prior, "m_col4_hier_prior.rds")

```

```{r}
m_col5_hier_prior <- brm(
  dplagoal ~ playersprints_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_col5,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
p_col5_hier <- pp_check(m_col5_hier_prior, ndraws = 100)
ggsave("col5_ppcheck_hier.png", p_col5_hier)
```

```{r}
m_col6_hier_prior <- brm(
  dplagoal ~ playersprints_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_col5,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
p_col6_hier <- pp_check(m_col5_hier_prior, ndraws = 100)
ggsave("col6_ppcheck_hier.png", p_col56hier)
```

```{r}
m_col7_hier_prior <- brm(
  goal_game ~ plaavspeed_game_c + dhomegame + rank_gap_c +
    (1 | player) + (1 | gamenu),
  data      = data_col7,
  family    = bernoulli(link = "logit"),
  prior     = priors_col7,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
p_col7_hier <- pp_check(m_col7_hier_prior, ndraws = 100)
ggsave("col7_ppcheck_hier.png", p_col7_hier)
```

## Posterior Check Hierachical Table 2 ----
```{r}
p_col2_hier <- pp_check(m_col2_hier, ndraws = 100)
ggsave("col2_posteriorcheck_hier.png", p_col2_hier)
```
```{r}
conditional_effects(
  m_col2_hier,                         
  effects   = "plaavspeed_c",           
  spaghetti = TRUE,
  ndraws    = 200,
  re_formula = NA                       
) %>%
  plot(
    points = FALSE,
    point_args = list(alpha = 1/2, size = 1),
    line_args  = list(colour = "hotpink")
  )
```

# Hierarchical models Table 2 ----
```{r}
# ---- HIERARCHICAL / MULTILEVEL (Table 2), cmdstanr backend, save model + LOO/WAIC ----

# ---------- Column 1 ----------
m_col1_hier <- brm(
  dplagoal ~ plaavspeed_c + (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col1,
  family    = bernoulli(link = "logit"),
  prior     = priors_col1,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col1_hier <- add_criterion(m_col1_hier, c("loo", "waic"))
saveRDS(m_col1_hier, "m_col1_hier.rds")

# ---------- Column 2 ----------
m_col2_hier <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col2,
  family    = bernoulli(link = "logit"),
  prior     = priors_col2,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col2_hier <- add_criterion(m_col2_hier, c("loo", "waic"))
saveRDS(m_col2_hier, "m_col2_hier.rds")

# ---------- Column 3 ----------
m_col3_hier <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_col3,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col3_hier <- add_criterion(m_col3_hier, c("loo", "waic"))
saveRDS(m_col3_hier, "m_col3_hier.rds")

# ---------- Column 4 ----------
m_col4_hier <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    team_crowd_c + opp_crowd_c + p_odds_team_c + p_odds_draw_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col4,
  family    = bernoulli(link = "logit"),
  prior     = priors_col4,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col4_hier <- add_criterion(m_col4_hier, c("loo", "waic"))
saveRDS(m_col4_hier, "m_col4_hier.rds")
```

```{r}

# ---------- Column 5 ----------
m_col5_hier <- brm(
  dplagoal ~ playersprints_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_col5,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col5_hier <- add_criterion(m_col5_hier, c("loo", "waic"))
saveRDS(m_col5_hier, "m_col5_hier.rds")

# ---------- Column 6 ----------
m_col6_hier <- brm(
  dplagoal ~ plaavspeed_c + playersprints_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_col6,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col6_hier <- add_criterion(m_col6_hier, c("loo", "waic"))
saveRDS(m_col6_hier, "m_col6_hier.rds")

# ---------- Column 7 ----------
m_col7_hier <- brm(
  goal_game ~ plaavspeed_game_c + dhomegame + rank_gap_c +
    (1 | player) + (1 | gamenu),
  data      = data_col7,
  family    = bernoulli(link = "logit"),
  prior     = priors_col7,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col7_hier <- add_criterion(m_col7_hier, c("loo", "waic"))
saveRDS(m_col7_hier, "m_col7_hier.rds")

```



# Intercept Only Table 2 ----
```{r}
# ---- Intercept-only (Table 2 baseline), cmdstanr backend, save + LOO/WAIC ----
threads_per_chain <- 2
Sys.setenv(STAN_NUM_THREADS = threads_per_chain)

m_intercept <- brm(
  dplagoal ~ 1,
  data      = data_base,
  family    = bernoulli(link = "logit"),
  prior     = set_prior("normal(-4.6, 0.3)", class = "Intercept"),
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

m_intercept <- add_criterion(m_intercept, c("loo", "waic"))
saveRDS(m_intercept, "m_intercept.rds")
```


# Looser Priors Table 2 ----
```{r}
# ---- PRIOR SENSITIVITY (LOOSER PRIORS) for Table 2: Columns 1 and 6 (Fixed Effects) ----
threads_per_chain <- 2
Sys.setenv(STAN_NUM_THREADS = threads_per_chain)

# looser FE base priors (applies to all factor() dummies + slopes unless overridden)
priors_fixef_base_loose <- c(
  set_prior("normal(-4.6, 0.6)", class = "Intercept"),
  set_prior("normal(0, 0.8)",    class = "b")
)

# ---- Column 1: dplagoal ~ plaavspeed_c + FE ----
priors_fixef_col1_loose <- c(
  priors_fixef_base_loose,
  set_prior("normal(0, 0.8)", class = "b", coef = "plaavspeed_c")
)

m_col1_loose <- brm(
  dplagoal ~ plaavspeed_c + factor(player) + factor(gamenu) + factor(section),
  data      = data_col1,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col1_loose,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col1_loose <- add_criterion(m_col1_loose, c("loo", "waic"))
saveRDS(m_col1_loose, "m_col1_loose.rds")

# ---- Column 6: dplagoal ~ plaavspeed_c + playersprints_c + controls + FE ----

# ---- PRIOR SENSITIVITY (LOOSER PRIORS) for Table 2: Columns 1 and 6 (Fixed Effects) ----

priors_fixef_col6_loose <- c(
  priors_fixef_base_loose,
  set_prior("normal(0, 0.8)", class = "b", coef = "plaavspeed_c"),
  set_prior("normal(0, 0.8)", class = "b", coef = "playersprints_c"),
  set_prior("normal(0, 1.0)", class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.3)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.2)", class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.1)", class = "b", coef = "player_cumsections_2_c"),
  set_prior("normal(0, 0.3)", class = "b", coef = "scoregap_pre_c"),
  set_prior("normal(0, 1.0)", class = "b", coef = "dteamgoal_pre"),
  set_prior("normal(0, 1.0)", class = "b", coef = "dplagoal_pre"),
  set_prior("normal(0, 1.0)", class = "b", coef = "dteamred_pre"),
  set_prior("normal(0, 0.5)", class = "b", coef = "dteaminj_sec_pre"),
  set_prior("normal(0, 0.5)", class = "b", coef = "team_avespeed_pre_c")
)

m_col6_loose <- brm(
  dplagoal ~ plaavspeed_c + playersprints_c +
    dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre +
    team_avespeed_pre_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col6_loose,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col6_loose <- add_criterion(m_col6_loose, c("loo", "waic"))
saveRDS(m_col6_loose, "m_col6_loose.rds")

```


# Evaluation Table 2 ----
## Results Fixed effects
```{r}
#summarize and print the results for all models in terms of probabilities

inv_logit <- function(x) 1 / (1 + exp(-x))

# summarise a numeric vector of draws
summ_num <- function(x) {
  data.frame(
    mean  = mean(x),
    sd    = sd(x),
    q2.5  = unname(quantile(x, 0.025)),
    q97.5 = unname(quantile(x, 0.975))
  )
}

# main function: returns a small table with up to 3 rows
# - Intercept is transformed to baseline probability p0
# - slopes are transformed to Δp for a +1 unit change in the covariate
show_three_prob <- function(m) {

  draws <- as_draws_df(m)

  out <- list()

  # 1) Intercept -> baseline probability
  alpha <- draws$b_Intercept
  p0 <- inv_logit(alpha)
  out[["Intercept (baseline p)"]] <- summ_num(p0)

  # 2) plaavspeed_c -> change in probability for +1 unit
  if ("b_plaavspeed_c" %in% variables(draws)) {
    beta <- draws$b_plaavspeed_c
    dp <- inv_logit(alpha + beta) - inv_logit(alpha)
    out[["plaavspeed_c (Δp for +1)"]] <- summ_num(dp)
  }

  # 3) playersprints_c -> change in probability for +1 unit
  if ("b_playersprints_c" %in% variables(draws)) {
    beta <- draws$b_playersprints_c
    dp <- inv_logit(alpha + beta) - inv_logit(alpha)
    out[["playersprints_c (Δp for +1)"]] <- summ_num(dp)
  }

  # bind into one compact table
  res <- do.call(rbind, out)
  res$variable <- rownames(res)
  rownames(res) <- NULL
  res[, c("variable", "mean", "sd", "q2.5", "q97.5")]
}

# ============================================================
# Run for all your models in one go
# ============================================================

models <- list(
  m_intercept  = m_intercept,
  m_col1       = m_col1,
  m_col2       = m_col2,
  m_col3       = m_col3,
  m_col4       = m_col4,
  m_col5       = m_col5,
  m_col6       = m_col6,
  m_col7       = m_col7,
  m_col1_loose = m_col1_loose,
  m_col6_loose = m_col6_loose
)

results_prob <- lapply(models, show_three_prob)

# print all results
for (nm in names(results_prob)) {
  cat("\n====================", nm, "====================\n")
  print(results_prob[[nm]])
}

```


## Results Hierachical ----
```{r}
#summarize and print the results for all models in terms of probabilities

inv_logit <- function(x) 1 / (1 + exp(-x))

# summarise a numeric vector of draws
summ_num <- function(x) {
  data.frame(
    mean  = mean(x),
    sd    = sd(x),
    q2.5  = unname(quantile(x, 0.025)),
    q97.5 = unname(quantile(x, 0.975))
  )
}

# main function: returns a small table with up to 3 rows
# - Intercept is transformed to baseline probability p0
# - slopes are transformed to Δp for a +1 unit change in the covariate
show_three_prob <- function(m) {

  draws <- as_draws_df(m)

  out <- list()

  # 1) Intercept -> baseline probability
  alpha <- draws$b_Intercept
  p0 <- inv_logit(alpha)
  out[["Intercept (baseline p)"]] <- summ_num(p0)

  # 2) plaavspeed_c -> change in probability for +1 unit
  if ("b_plaavspeed_c" %in% variables(draws)) {
    beta <- draws$b_plaavspeed_c
    dp <- inv_logit(alpha + beta) - inv_logit(alpha)
    out[["plaavspeed_c (Δp for +1)"]] <- summ_num(dp)
  }

  # 3) playersprints_c -> change in probability for +1 unit
  if ("b_playersprints_c" %in% variables(draws)) {
    beta <- draws$b_playersprints_c
    dp <- inv_logit(alpha + beta) - inv_logit(alpha)
    out[["playersprints_c (Δp for +1)"]] <- summ_num(dp)
  }

  # bind into one compact table
  res <- do.call(rbind, out)
  res$variable <- rownames(res)
  rownames(res) <- NULL
  res[, c("variable", "mean", "sd", "q2.5", "q97.5")]
}

# ============================================================
# Run for all your models in one go
# ============================================================

models <- list(
  m_col1_hier       = m_col1_hier,
  m_col2_hier       = m_col2_hier,
  m_col3_hier       = m_col3_hier,
  m_col4_hier       = m_col4_hier,
  m_col5_hier       = m_col5_hier,
  m_col6_hier       = m_col6_hier,
  m_col7_hier       = m_col7_hier
)

results_prob <- lapply(models, show_three_prob)

# print all results
for (nm in names(results_prob)) {
  cat("\n====================", nm, "====================\n")
  print(results_prob[[nm]])
}

```

## Comparison LOO
```{r}
#Fixed effects
loo_compare(m_intercept, m_col1, m_col2)

loo_compare(m_col3, m_col5, m_col6)
```

```{r}
#Hierarchical effects
loo_compare(m_intercept, m_col1_hier, m_col2_hier)

loo_compare(m_col3_hier, m_col5_hier, m_col6_hier)
```

```{r}
#Comparison
loo_compare(m_col1, m_col1_hier)

loo_compare(m_col2, m_col2_hier)

loo_compare(m_col3, m_col3_hier)

loo_compare(m_col4, m_col4_hier)

loo_compare(m_col5, m_col5_hier)

loo_compare(m_col6, m_col6_hier)

loo_compare(m_col7, m_col7_hier)
```

## Evaluation R-hat
```{r}
# evaluation r-hat fixed effects
models <- list(
  m_intercept,
  m_col1, m_col2, m_col3, m_col4, m_col5, m_col6, m_col7,
  m_col1_loose, m_col6_loose, m_intercept
)

sapply(models, function(m) max(rhat(m), na.rm = TRUE))
```

```{r}
# evaluation r-hat hierachical
models <- list(
  m_col1_hier, m_col2_hier, m_col3_hier, m_col4_hier, m_col5_hier, m_col6_hier, m_col7_hier
)

sapply(models, function(m) max(rhat(m), na.rm = TRUE))
```


# Table 3


## Add Team Names for table 3
```{r}
# add team names to the data for table 3
library(readxl)
team_names <- read_excel("TeamID.xlsx")


#teams don't play the same amount of games, there are 26 games in the first round then they are divided in top teams and bottom teams. isolate the 26 first games (Appendix A.2, in the paper) the regular season



# team rank in game 26 and game 27 because rank variable is the rank before the game

team_rank_r26 <- data %>%
  filter(played == 26) %>%           # only round 26
  distinct(team, team_rank)          # one row per team with its rank

team_rank_r27 <- data %>%
  filter(played == 27) %>%           # only round 27
  distinct(team, team_rank)          # one row per team with its rank

# compared with table "Positions by round" from https://en.wikipedia.org/wiki/2017%E2%80%9318_Israeli_Premier_League#Regular_season_table  -> team_rank_r27 is the right one

# add the israeli name to round 27
team_rank_r27 <- team_rank_r27 %>%
  left_join(
    team_names %>%
      select(team_rank, israeli_name = `händisch israeli name`),
    by = "team_rank"
  )

#add the israeli name to main data
data <- data %>%
  left_join(
    team_rank_r27 %>%
      select(team, israeli_name),
    by = "team"
  )
```


```{r}
##### First, aggregate on the team x game level and keep variables of interest #####

data_teamxgame <- data %>% 
  group_by(gamenu, team, opponent) %>%  # we keep opponent to include pair fixed effects
  summarise(
    teamdist_game     = unique(teamdist_game),
    oppdist_game      = unique(oppdist_game),
    dteamwin          = unique(dteamwin),     # same for all players on team
    team_crowd        = unique(team_crowd),
    opp_crowd         = unique(opp_crowd),
    rank_gap          = unique(rank_gap),
    dhomegame         = unique(dhomegame),
    danyteamredc_game = unique(danyteamredc_game),
    israeli_name      = unique(israeli_name),
    played            = unique(played),
    p_odds_team       = unique(p_odds_team),
    p_odds_draw       = unique(p_odds_draw),
    .groups = "drop"
  ) %>% 
  mutate(
    distance_ratio = teamdist_game / oppdist_game,
    total_distance = teamdist_game + oppdist_game,
    # create pair fixed effects before dropping one team per game
    team_low  = pmin(team, opponent),
    team_high = pmax(team, opponent),
    pair      =paste(team_low, team_high, sep = "_")
  ) %>%
  group_by(gamenu) %>%
  slice_min(israeli_name) %>%      # keep only the team which comes first in the hebrew alpahbet
  ungroup()
```

pair fixed effects
```{r}
data_teamxgame <- data_teamxgame %>%
  mutate(
    team_low  = pmin(team, opponent),
    team_high = pmax(team, opponent),
    pair = paste0(team_low, "_", team_high)
  )
# we have 14 teams, 14 over 2 = 91 possible matchups, verify below
length(unique(data_teamxgame$pair))
```


```{r}
# ---------- Create derived variables and center continuous regressors ----------
data_teamxgame <- data_teamxgame %>%
  mutate(
    distance_ratio    = ifelse(is.na(distance_ratio), teamdist_game / ifelse(oppdist_game==0, NA, opp_dist),     distance_ratio),
    total_distance    = teamdist_game + oppdist_game,
    # center continuous predictors (mean-centering)
    teamdist_c        = teamdist_game - mean(teamdist_game, na.rm = TRUE),
    oppdist_c         = oppdist_game  - mean(oppdist_game,  na.rm = TRUE),
    distance_ratio_c  = distance_ratio - mean(distance_ratio, na.rm = TRUE),
    total_distance_c  = total_distance - mean(total_distance, na.rm = TRUE),
    rank_gap_c        = rank_gap - mean(rank_gap, na.rm = TRUE),
    team_crowd_c      = team_crowd - mean(team_crowd, na.rm = TRUE),
    opp_crowd_c       = opp_crowd - mean(opp_crowd, na.rm = TRUE),
    p_odds_team_c     = p_odds_team - mean(p_odds_team, na.rm = TRUE),
    p_odds_draw_c     = p_odds_draw - mean(p_odds_draw, na.rm = TRUE)
  )

# datasets for each column 
table3_col1 <- data_teamxgame %>%
  drop_na(dteamwin, teamdist_game, oppdist_game)  

table3_col2 <- data_teamxgame %>%
  drop_na(dteamwin, teamdist_game, oppdist_game, dhomegame, rank_gap)

table3_col3 <- data_teamxgame %>%
  drop_na(dteamwin, teamdist_game, oppdist_game, dhomegame, rank_gap, team_crowd_c, opp_crowd_c, p_odds_team_c, p_odds_draw_c)

table3_col4 <- data_teamxgame %>%
  # restrict to games without red cards
  filter(is.na(danyteamredc_game) | danyteamredc_game == 0) %>%
  drop_na(dteamwin, teamdist_game, oppdist_game, dhomegame, rank_gap)

table3_col5 <- data_teamxgame %>%
  drop_na(dteamwin, distance_ratio, dhomegame, rank_gap)

table3_col6 <- data_teamxgame %>%
  drop_na(dteamwin, distance_ratio, total_distance, dhomegame, rank_gap)

# double-check sample sizes
sapply(list(table3_col1, table3_col2, table3_col3, table3_col4, table3_col5, table3_col6), nrow)
```

slightly different sample sizes compared to the ones from table 3 in the paper (why?)


# Intercept-only
```{r}
m_t3_intercept <- brm(
  dteamwin ~ 1, 
  data    = data_teamxgame,
  family  = bernoulli(link = "logit"),
  sample_prior = "only",
  prior = prior( normal(0, 1), class = Intercept),      
  iter    = 1000, warmup = 500, chains = 4, cores = 4
)

pp_check(m_t3_intercept)

#saveRDS(m_t3_intercept, "m_t3_intercept.rds")
```

```{r}
readRDS("m_t3_intercept.rds")
mean(data_teamxgame$dteamwin)
```
Intercept estimate: -0.48 --> 38.2%
Mean also ~38%


#col1
```{r}
#set weakly informative priors that fit logit scale 
priors_t3c1 <- c( 
  set_prior("normal(-0.4, 1)", class = "Intercept"), 
  set_prior("normal(0, 0.5)", class = "b", coef = "teamdist_c"),
  set_prior("normal(0, 0.5)", class = "b", coef = "oppdist_c"),
  # Multilevel regularization
  set_prior("exponential(1)", class = "sd", group = "pair"),
  set_prior("exponential(1)", class = "sd", group = "played")
)
```


```{r}
m_t3col1_prior <- brm( 
  dteamwin ~ teamdist_c + oppdist_c + (1 | pair) + (1 | played),
  data = table3_col1, 
  family = bernoulli(link = "logit"),
  sample_prior = "only",
  prior = priors_t3c1,
  iter = 4000, warmup = 2000, chains = 4, cores = 4,
  control = list(max_treedepth = 15, adapt_delta = 0.95) #increase treedepth to avoid problems during warmup
)
```

```{r}
#prior predictive check
pp_check(m_t3col1_prior, ndraws = 100)          
yrep_small <- posterior_predict(m_t3col1_prior, ndraws = 100)
hist(rowMeans(yrep_small))
plot(m_t3col1_prior, N = 1000)                             
```

```{r}
m_t3col1 <- brm(
  dteamwin ~ teamdist_c + oppdist_c + (1 | pair) + (1 | played),
  data    = table3_col1,
  family  = bernoulli(link = "logit"),
  prior   = priors_t3c1,
  iter    = 4000, warmup = 2000, chains = 4, cores = 4
)
print(summary(m_t3col1), digits = 6)
```

```{r}
summary(m_t3col1)
pp_check(m_t3col1, ndraws = 100)
#conditional_effects(m_t3col1, spaghetti = TRUE, ndraws = 200)
plot(m_t3col1)
```

```{r}
cor(table3_col1 %>% select(teamdist_c, oppdist_c, distance_ratio_c, total_distance_c), use = "pairwise.complete.obs")
```


#col2
```{r}
priors_t3c2 <- c( 
  set_prior("normal(-0.4, 1)", class = "Intercept"),
  set_prior("normal(0, 0.5)", class = "b", coef = "teamdist_c"),
  set_prior("normal(0, 0.5)", class = "b", coef = "oppdist_c"), 
  set_prior("normal(0, 0.5)",  class = "b", coef = "dhomegame"), #reasonable prior (expect small effects), otherwise set sd=1 if we expect stronger effects 
  set_prior("normal(0, 0.3)", class = "b", coef = "rank_gap_c"),
  set_prior("exponential(1)", class = "sd", group = "pair"),
  set_prior("exponential(1)", class = "sd", group = "played")
)
```

```{r}
m_t3col2 <- brm(
  dteamwin ~ teamdist_c + oppdist_c + (1 | pair) + (1 | played) +
    dhomegame +
    rank_gap_c,     
  data    = table3_col2,
  #sample_prior = "only",
  family  = bernoulli(link = "logit"),
  prior   = priors_t3c2,
  iter    = 4000, warmup = 2000, chains = 4, cores = 4,
  control = list(max_treedepth = 15, adapt_delta = 0.95) #increase treedepth to avoid problems during warmup
)
print(summary(m_t3col2), digits = 6)
#plot(m_t3col2)
```

```{r}
pp_check(m_t3col2, ndraws = 100)
```


#col3
```{r}
priors_t3c3 <- c( 
  set_prior("normal(-0.4, 1)", class = "Intercept"),
  set_prior("normal(0, 0.5)", class = "b", coef = "teamdist_c"),     #
  set_prior("normal(0, 0.5)", class = "b", coef = "oppdist_c"),      #
  set_prior("normal(0, 0.5)",  class = "b", coef = "dhomegame"),     #
  set_prior("normal(0, 0.3)", class = "b", coef = "rank_gap_c"),     #
  set_prior("normal(0, 0.5)", class = "b", coef = "team_crowd_c"),   #
  set_prior("normal(0, 0.5)", class = "b", coef = "opp_crowd_c"),    #
  set_prior("normal(0, 1)", class = "b", coef = "p_odds_team_c"),    # sensitive priors
  set_prior("normal(0, 1)", class = "b", coef = "p_odds_draw_c"),    # sensitive priors
  set_prior("exponential(1)", class = "sd", group = "pair"),         #
  set_prior("exponential(1)", class = "sd", group = "played")        #
)
```

```{r}
m_t3col3 <- brm(
  dteamwin ~ 
    teamdist_c + 
    oppdist_c + 
    team_crowd_c +
    opp_crowd_c +
    p_odds_team_c +
    p_odds_draw_c +
    (1 | pair) + 
    (1 | played) +
    dhomegame +
    rank_gap_c,     
  data    = table3_col3,
  #sample_prior = "only",
  family  = bernoulli(link = "logit"),
  prior   = priors_t3c3,
  iter    = 4000, warmup = 2000, chains = 4, cores = 4,
  control = list(max_treedepth = 15, adapt_delta = 0.95) #increase treedepth to avoid problems during warmup
)
print(summary(m_t3col3), digits = 6)
#plot(m_t3col3)
```

```{r}
pp_check(m_t3col3, ndraws = 100)
```

#col4
```{r}
priors_t3c4 <- c( 
  set_prior("normal(-0.4, 1)", class = "Intercept"),
  set_prior("normal(0, 0.5)", class = "b", coef = "teamdist_c"), #
  set_prior("normal(0, 0.5)", class = "b", coef = "oppdist_c"),  #
  set_prior("normal(0, 0.5)",  class = "b", coef = "dhomegame"), #
  set_prior("normal(0, 0.5)", class = "b", coef = "rank_gap_c"), #
  set_prior("exponential(1)", class = "sd", group = "pair"),     #
  set_prior("exponential(1)", class = "sd", group = "played")    #
)
```

```{r}
m_t3col4 <- brm(
  dteamwin ~ teamdist_c + oppdist_c + (1 | pair) + (1 | played) +
    dhomegame +
    rank_gap_c,     
  data    = table3_col4,
  #sample_prior = "only",
  family  = bernoulli(link = "logit"),
  prior   = priors_t3c4,
  iter    = 4000, warmup = 2000, chains = 4, cores = 4,
  control = list(max_treedepth = 15, adapt_delta = 0.95) #increase treedepth to avoid problems during warmup
)
print(summary(m_t3col4), digits = 6)
#plot(m_t3col4)
```

```{r}
pp_check(m_t3col4, ndraws = 100)
```


#col5

below we impose reasonably skeptical priors. What the priors say:
Distance ratio matters, but not absurdly.
Home advantage is modest.
Rank gap matters a little, but not too much.

```{r}
priors_t3c5 <-c(
  set_prior("normal(-0.4, 1)", class = "Intercept"),  
  set_prior("normal(0, 0.5)",  class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.3)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 2)", class = "b", coef = "distance_ratio_c"), #sensitive prior
  set_prior("exponential(1)", class = "sd", group = "pair"),
  set_prior("exponential(1)", class = "sd", group = "played")
)
```

```{r}
m_t3col5 <- brm(
  dteamwin ~ distance_ratio_c + (1 | pair) + (1 | played) +
    dhomegame +
    rank_gap_c,     
  data    = table3_col5,
  #sample_prior = "only",
  family  = bernoulli(link = "logit"),
  prior   = priors_t3c5,
  iter    = 4000, warmup = 2000, chains = 4, cores = 4,
  control = list(max_treedepth = 15, adapt_delta = 0.95) #increase treedepth to avoid problems during warmup
)
print(summary(m_t3col5), digits = 6)
```

```{r}
pp_check(m_t3col5, ndraws = 100)
```


#col6
how do we choose priors?
distance_ratio --> relative effort (could have moderate effect)
total_distance --> game intensity/pace (probably weak effect if at all? guessing its near zero)
aim for weak regularization

```{r}
priors_t3c6 <-c(
   set_prior("normal(-0.4, 1)", class = "Intercept"),  
  set_prior("normal(0, 0.5)",  class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.3)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 2)", class = "b", coef = "distance_ratio_c"), #very prior sensitive
  set_prior("normal(0, 5)", class = "b", coef = "total_distance_c"),
  set_prior("exponential(1)", class = "sd", group = "pair"),
  set_prior("exponential(1)", class = "sd", group = "played")
)
```

```{r}
m_t3col6 <- brm(
  dteamwin ~ distance_ratio_c + total_distance_c + (1 | pair) + (1 | played) +
    dhomegame +
    rank_gap_c,     
  data    = table3_col6,
  #sample_prior = "only",
  family  = bernoulli(link = "logit"),
  prior   = priors_t3c6,
  iter    = 4000, warmup = 2000, chains = 4, cores = 4,
  control = list(max_treedepth = 15, adapt_delta = 0.95) #increase treedepth to avoid problems during warmup
)
print(summary(m_t3col6), digits = 6)
#plot(m_t3col6)
```

```{r}
pp_check(m_t3col6, ndraws = 100)
```


#loo evaluation for table 3
```{r}
library(loo)
# loo requires same subsamples, however we have 2 colums (3 and 4) which are estimated on different data sets, therefore they are excluded in the comparison
loo_t3c1 <- loo(m_t3col1)
loo_t3c2 <- loo(m_t3col2)
loo_t3c3 <- loo(m_t3col3)
loo_t3c4 <- loo(m_t3col4)
loo_t3c5 <- loo(m_t3col5)
loo_t3c6 <- loo(m_t3col6)

# compare models
loo_compare(
  loo_t3c1,
  loo_t3c2,
 # loo_t3c3, 180 observations
 # loo_t3c4, 152 observations
  loo_t3c5,
  loo_t3c6
)
```



# multicollinearity plot
```{r}
# table 3: multicollinearity plot


library(brms)
library(tidyverse)

# m_t3c4 = your fitted brms model that used priors_t3c4
post <- as_draws_df(m_t3col4) %>%
  select(b_teamdist_c, b_oppdist_c)

ggplot(post, aes(x = b_teamdist_c, y = b_oppdist_c)) +
  geom_point(alpha = 0.12, size = 0.8) +
  geom_density_2d() +
  labs(
    title = "Posterior ridge: teamdist_c vs oppdist_c",
    x = expression(beta[teamdist_c]),
    y = expression(beta[oppdist_c])
  ) +
  theme_minimal()



ols <- lm(dteamwin ~ teamdist_c + oppdist_c + dhomegame + rank_gap_c,
          data = table3_col4)

ols_pt <- tibble(
  b_teamdist_c = coef(ols)[["teamdist_c"]],
  b_oppdist_c  = coef(ols)[["oppdist_c"]]
)

ggplot(post, aes(x = b_teamdist_c, y = b_oppdist_c)) +
  geom_point(alpha = 0.12, size = 0.8) +
  geom_density_2d() +
  geom_point(data = ols_pt, aes(x = b_teamdist_c, y = b_oppdist_c),
             size = 4) +
  labs(
    title = "Posterior ridge with OLS point",
    x = expression(beta[teamdist_c]),
    y = expression(beta[oppdist_c])
  ) +
  theme_minimal()




post %>%
  mutate(beta_diff = b_teamdist_c - b_oppdist_c) %>%
  ggplot(aes(x = beta_diff)) +
  geom_density(fill = "grey80") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    title = "Identified effect direction",
    x = expression(beta[teamdist_c] - beta[oppdist_c])
  ) +
  theme_minimal()

post2 <- post %>%
  mutate(
    beta_sum  = b_teamdist_c + b_oppdist_c,
    beta_diff = b_teamdist_c - b_oppdist_c
  )

ggplot(post2, aes(beta_sum)) +
  geom_density(fill = "grey85") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(title = "Identified combination: beta_teamdist + beta_oppdist",
       x = expression(beta[teamdist] + beta[oppdist])) +
  theme_minimal()


library(tidybayes)
library(dplyr)
library(ggplot2)

draws_dr <- m_t3col5 %>%
  spread_draws(b_distance_ratio_c) %>%
  rename(beta = b_distance_ratio_c)

p_pos <- mean(draws_dr$beta > 0)

# compute ONE density curve
d <- density(draws_dr$beta, n = 2048)
dens_df <- tibble(
  x = d$x,
  y = d$y,
  above0 = d$x > 0
)

ggplot(dens_df, aes(x = x, y = y)) +
  # fill left side (<=0)
  geom_area(data = dens_df %>% filter(!above0),
            fill = "grey85", alpha = 1) +
  # fill right side (>0)
  geom_area(data = dens_df %>% filter(above0),
            fill = "grey55", alpha = 1) +
  # outline of the same density curve
  geom_line(color = "grey30", linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    title = paste0(
      "Posterior for distance_ratio_c:  P(β > 0) = ",
      sprintf("%.3f", p_pos)
    ),
    x = expression(beta[distance_ratio_c]),
    y = "Density"
  ) +
  theme_minimal()



```
