---
title: "Replication Challenge"
author: "Lisa Philine Krämer"
date: "2025-11-20"
output: html_document
---
# Setup ----
```{r}
library(tidyverse)
library(haven)
library(ggplot2)
library(dplyr)
library(plotly)
library(tidyr)
library(xtable)
library(knitr)
library(kableExtra)
library(tidybayes)
library(brms)
library(bayesplot)
library(posterior)
library(dagitty)
library(ggdag)
library(cmdstanr)
library(brms)
library(loo)
library(brms)
library(bayesplot)

dir.create("ppc_prior_hier", showWarnings = FALSE)

```

```{r}
#Lourenco: Was ist das?
#mean(is.na(data$opp_crowd)) * 100
```


```{r}
data <- read_dta("original&computed_DB_forrep.dta")
```

# Data Preparation and Exploration ----
```{r}
# sort the columns to be alphabetically ordered
data <- data[, order(names(data))]

# change dates to Date type 
data$bdate <- as.Date(data$bdate, format = "%Y-%m-%d")
data$joineddate <- as.Date(data$joineddate, format = "%Y-%m-%d")
```

```{r}
summary(data)
```

```{r}
data %>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
  facet_wrap(~ key, scales = "free") +   # In separate panels
  geom_histogram() 
```



```{r}
# Count of NAs per column
colSums(is.na(data))

# Percentage of missing values
colSums(is.na(data)) / nrow(data) * 100

```

```{r}
#plot every variable's histogram, since all data is either numeric or date

all_vars <- names(data)

for (v in all_vars) {
  print(
    ggplot(data, aes_string(x = v)) +
      geom_histogram(bins = 30) +
      ggtitle(paste("Histogram of", v))
  )
}


```

```{r}
# heatmap


# make a numeric-only version of your data:
data_num <- data %>%
  mutate(across(
    where(~ inherits(.x, "Date")),  # <- predicate: is it a Date?
    as.numeric
  ))

# calculate correlations
cor_mat <- cor(data_num, use = "pairwise.complete.obs", method = "pearson")

library(plotly)

# custom red–white–blue colorscale: blue (negative) → white (0) → red (positive)
rwb_scale <- list(
  list(0.0, "blue"),
  list(0.5, "white"),
  list(1.0, "red")
)

plot_ly(
  x = colnames(cor_mat),
  y = rownames(cor_mat),
  z = cor_mat,
  type = "heatmap",
  zmin = -1,
  zmax = 1,
  colorscale = rwb_scale,
  hovertemplate = paste(
    "Var1: %{x}<br>",
    "Var2: %{y}<br>",
    "Correlation: %{z:.2f}<extra></extra>"
  )
) %>%
  layout(
    title = list(
      text = "Pearson Correlation Heatmap",
      x = 0.5,
      xanchor = "center"
    ),
    xaxis = list(
      title = "",
      tickangle = -45,
      tickfont = list(size = 9)
    ),
    yaxis = list(
      title = "",
      tickfont = list(size = 9,
                      automargin = TRUE)
    ),
    margin = list(l = 80, r = 20, b = 120, t = 50),
    coloraxis = list(
      colorbar = list(
        title = "r",
        ticksuffix = "",
        len = 0.8
      )
    )
  )


```






  1##Prepare Data Sets
```{r}
data_base <- data %>%
  filter(
    dfullsection == 1,          # full sections only
    position %in% c(3, 4),      # midfielders + strikers
    !(section %in% c(10, 20)) 
  )
data_base <- data_base  %>% mutate(player = factor(player),
                                   gamenu = factor(gamenu),
                                   section = factor(section))
```

## DAGs
```{r}
dag1 <- dagitty("dag{
                S -> G
                NoS -> G
                NoS -> S 
                E -> S
                E -> NoS}")

coordinates(dag1) <- list( 
                          x = c(S = 0, NoS = 1, G = 2, E = 0) , 
                          y = c(S = 0, NoS = 1, G = 0, E = 1))

ggdag( dag1 ) + 
  ggtitle( "(1) Effort influences Speed & Number of Sprints;\n(2) Speed influences Goal Probability;\n(3) Number of Sprints influences Speed & Goal Probability" ) +
  theme_dag()



dag2 <- dagitty("dag{
                D -> W
                OD -> W
                }")

coordinates(dag2) <- list( 
                          x = c(D = 0, OD = 2, W = 1) , 
                          y = c(D = 0, OD = 0, W = 0))

ggdag( dag2 ) + 
  ggtitle( "(1) Team Distance influences Team Winning Probability;\n(2) Opponent Distance influences Team Winning Probability" ) +
  theme_dag()
```
```{r}
library(fixest)

reg1 <- feols(dplagoal ~ plaavspeed, data = data_base)
reg2 <- feols(dplagoal ~ playersprints, data = data_base)
reg3 <- feols(dplagoal ~ plaavspeed + playersprints, data = data_base)
reg4 <- feols(plaavspeed ~ playersprints, data = data_base)

etable(reg1, reg2, reg3, reg4)

```

```{r}
data %>%
  ggplot(aes(x=playersprints, y=plaavspeed)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  theme_classic()

data %>%
  ggplot(aes(x=teamdist_game, y = oppdist_game)) +
  geom_point() +
  geom_smooth(method = "lm", se = T) +
  theme_classic()


data %>%
  ggplot(aes(x=dplagoal)) +
  geom_histogram() +
  theme_classic()


data %>%
  ggplot(aes(x=dteamwin)) +
  geom_histogram() +
  theme_classic()
```




## Summary Statistics

Define missing variable

```{r}
data_base <- data_base %>%
  mutate(
    SMGP = ifelse(position %in% c(3,4),          # SMGP = Strikers and midfielders goal probability
                  dplagoal,                      # already 0/1 if player scored
                  NA)                            # set others to NA
  )
```

```{r}
# unique combinations of team and games
TG_combinations <- data_base %>%
  distinct(team, gamenu) %>%
  nrow()

TG_combinations
```

```{r}
team_game_redcards <- data_base %>%
  group_by(team, gamenu) %>%
  summarise(
    redcard_teamgame = as.integer(any(dplaredc == 1)),
    .groups = "drop"
  )

```

```{r}
mean(data$dfullsection, na.rm = TRUE)
```


```{r}
### --------------------------------------------------------------
### PANEL A: Section level
### --------------------------------------------------------------

panelA_vars <- list(
  full_sections                = "dfullsection",
  overtime_sections            = "(data$section %in% c(10, 20)),2)",
  player_distance              = "playerdist",
  player_sprints               = "playersprints",
  goal_probability             = "SMGP",
  team_distance                = "teamdist"
)

panelA <- data_base %>% 
  summarise(
    FullSections_mean     = mean(data$dfullsection, na.rm = TRUE),
    Overtime_mean         = round(mean(data$section %in% c(10, 20)),2),
    PlayerDistance_mean   = mean(playerdist, na.rm = TRUE),
    PlayerDistance_sd     = sd(playerdist, na.rm = TRUE),
    PlayerSprints_mean    = mean(playersprints, na.rm = TRUE),
    PlayerSprints_sd      = sd(playersprints, na.rm = TRUE),
    GoalProb_mean         = mean(SMGP, na.rm = TRUE),
    GoalProb_sd           = sd(SMGP, na.rm = TRUE),
    TeamDistance_mean     = mean(teamdist, na.rm = TRUE),
    TeamDistance_sd       = sd(teamdist, na.rm = TRUE),
    Observations_PS       = nrow(data),
    Observations_TS       = TG_combinations*20         #18 oder 20?
  )

options(scipen = 999)
panelA <- panelA %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )

```

```{r}


dataB <- data %>%
  filter(dfullsection == 1 & !(section %in% c(10))) %>%
  group_by(gamenu, player) %>%
  summarise(
    player_sectionsplayed = first(player_sectionsplayed),  
    player_avg_speed      = mean(plaavspeed_game, na.rm = TRUE),
    .groups = "drop"
  )

dataB2 <- data %>%
  filter(dfullsection == 1 & !(section %in% c(10))) %>%
  group_by(gamenu, team) %>%
  summarise(
    teamdist_game = mean(teamdist_game, na.rm = TRUE)
  )

crowd_game <- data %>%
  group_by(gamenu, team, dhomegame) %>%  
  summarise(
    home_crowd = first(team_crowd),
    opp_crowd  = first(opp_crowd),
    .groups = "drop"
  )

redcard <- data %>%
  group_by(gamenu) %>%
  summarise(
    any_red = max(dplaredc, na.rm = TRUE),  
    .groups = "drop"
  )



```



```{r}
### --------------------------------------------------------------
### PANEL B: Game level
### --------------------------------------------------------------

panelB <- data %>% 
  summarise(
    SectionsPlayed_mean   = mean(dataB$player_sectionsplayed, na.rm = TRUE),
    SectionsPlayed_sd     = sd(dataB$player_sectionsplayed, na.rm = TRUE),
    AvgSpeed_mean         = mean(dataB$player_avg_speed, na.rm = TRUE),
    AvgSpeed_sd           = sd(dataB$player_avg_speed, na.rm = TRUE),
    TeamDistanceGame_mean = mean(dataB2$teamdist_game, na.rm = TRUE),
    TeamDistanceGame_sd   = sd(dataB2$teamdist_game, na.rm = TRUE),
    HomeCrowd_mean        = mean(crowd_game$home_crowd[crowd_game$dhomegame == 1], na.rm = TRUE),
    HomeCrowd_sd          = sd(crowd_game$home_crowd[crowd_game$dhomegame == 1], na.rm = TRUE),
    AwayCrowd_mean        = mean(crowd_game$opp_crowd[crowd_game$dhomegame == 1], na.rm = TRUE),
    AwayCrowd_sd          = sd(crowd_game$opp_crowd[crowd_game$dhomegame == 1], na.rm = TRUE),
    RedCardProb_mean      = mean(redcard$any_red, na.rm = TRUE),
    RedCardProb_sd        = sd(redcard$any_red, na.rm = TRUE),
    Obs_player_game       = data %>% 
      filter(dfullsection == 1 & !(section %in% c(10))) %>%
      distinct(gamenu, player) %>% nrow(),
    Obs_team_game         = n_distinct(team, gamenu),
    
    Rounds                = n_distinct(played),
    Games                 = n_distinct(gamenu),
    Teams                 = n_distinct(team),
    Players               = n_distinct(player),
    Strikers              = n_distinct(data$player[data$position == 4]),
    Midfielders           = n_distinct(data$player[data$position == 3]),
    Defenders             = n_distinct(data$player[data$position == 2])
  )

panelB <- panelB %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )
```

## Latex code for summary statistics
```{r}
### PANEL A Latex code
panelA_latex <- kable(panelA, 
                      format = "latex", 
                      booktabs = TRUE, 
                      caption = "Panel A: Section LeveL") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))

# Print Latex code
cat(panelA_latex)

### PANEL B Latex code
panelB_latex <- kable(panelB, 
                      format = "latex", 
                      booktabs = TRUE, 
                      caption = "Panel B: Game Level") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))

# Print Latex code
cat(panelB_latex)
```



# Data Preparation ----
```{r}
data_col1 <- data_base %>%
  drop_na(dplagoal, plaavspeed) %>%
  mutate(
    plaavspeed_c = plaavspeed - mean(plaavspeed, na.rm = TRUE)
  )
```

```{r}
data_col2 <- data_base %>%
  drop_na(
    dplagoal, plaavspeed,
    dhomegame, rank_gap,
    player_cumsections, player_cumsections_2
  ) %>%
  mutate(
    plaavspeed_c           = plaavspeed - mean(plaavspeed, na.rm = TRUE),
    rank_gap_c             = rank_gap - mean(rank_gap, na.rm = TRUE),
    player_cumsections_c   = player_cumsections - mean(player_cumsections, na.rm = TRUE),
    player_cumsections_2_c = player_cumsections_2 - mean(player_cumsections_2, na.rm = TRUE)
  )

```

```{r}
data_col356 <- data_base %>%
  drop_na(
    dplagoal, plaavspeed, playersprints,
    dhomegame, rank_gap,
    player_cumsections, player_cumsections_2,
    scoregap_pre, dteamgoal_pre, dplagoal_pre,
    dteamred_pre, dteaminj_sec_pre,
    team_avespeed_pre
  ) %>%
  mutate(
    plaavspeed_c           = plaavspeed - mean(plaavspeed, na.rm = TRUE),
    playersprints_c        = playersprints - mean(playersprints, na.rm = TRUE),
    rank_gap_c             = rank_gap - mean(rank_gap, na.rm = TRUE),
    player_cumsections_c   = player_cumsections - mean(player_cumsections, na.rm = TRUE),
    player_cumsections_2_c = player_cumsections_2 - mean(player_cumsections_2, na.rm = TRUE),
    scoregap_pre_c         = scoregap_pre - mean(scoregap_pre, na.rm = TRUE),
    team_avespeed_pre_c    = team_avespeed_pre - mean(team_avespeed_pre, na.rm = TRUE)
  )
```

```{r}
data_col4 <- data_base %>%
  drop_na(
    dplagoal, plaavspeed,
    dhomegame, rank_gap,
    player_cumsections, player_cumsections_2,
    scoregap_pre, dteamgoal_pre, dplagoal_pre,
    dteamred_pre, dteaminj_sec_pre,
    team_avespeed_pre,
    team_crowd, opp_crowd,
    p_odds_team, p_odds_draw
  ) %>%
  mutate(
    plaavspeed_c           = plaavspeed - mean(plaavspeed, na.rm = TRUE),
    rank_gap_c             = rank_gap - mean(rank_gap, na.rm = TRUE),
    player_cumsections_c   = player_cumsections - mean(player_cumsections, na.rm = TRUE),
    player_cumsections_2_c = player_cumsections_2 - mean(player_cumsections_2, na.rm = TRUE),
    scoregap_pre_c         = scoregap_pre - mean(scoregap_pre, na.rm = TRUE),
    team_avespeed_pre_c    = team_avespeed_pre - mean(team_avespeed_pre, na.rm = TRUE),
    team_crowd_c           = team_crowd - mean(team_crowd, na.rm = TRUE),
    opp_crowd_c            = opp_crowd - mean(opp_crowd, na.rm = TRUE),
    p_odds_team_c          = p_odds_team - mean(p_odds_team, na.rm = TRUE),
    p_odds_draw_c          = p_odds_draw - mean(p_odds_draw, na.rm = TRUE)
  )

```

```{r}
# player × game data for column 7 (as in the paper)
data_col7 <- data %>%
  # same sample restrictions as Table 2
  filter(
    position %in% c(3, 4),         # only midfielders and strikers
    !(section %in% c(10, 20))      # drop overtime sections
  ) %>%
  group_by(player, gamenu) %>%
  summarise(
    # outcome: at least one goal by this player in this game
    goal_game        = as.integer(any(dplagoal == 1)),
    
    # main regressor: game-average speed of the player
    plaavspeed_game  = first(plaavspeed_game),
    
    # controls that are constant within a game for that player
    dhomegame        = first(dhomegame),
    rank_gap         = first(rank_gap),
    
    .groups = "drop"
  ) %>%
  # drop missing info (e.g. if some game-level variable is NA)
  drop_na(goal_game, plaavspeed_game, dhomegame, rank_gap) %>%
  # centering (like in the other columns)
  mutate(
    plaavspeed_game_c = plaavspeed_game - mean(plaavspeed_game, na.rm = TRUE),
    rank_gap_c        = rank_gap - mean(rank_gap, na.rm = TRUE),
    # factors for player and game fixed/random effects
    player            = factor(player),
    gamenu            = factor(gamenu)
  )

```

```{r}
#check if n matches table
nrow(data_col1)
nrow(data_col2)
nrow(data_col356) #unterschieded sich um 2, müsste aber eig gleich sein
nrow(data_col4)
nrow(data_col7) #hier fehlen 14, verstehe nicht warum
```



# Setup Table 2 ----
```{r}
# ============================================================
# (A) HIERARCHICAL priors (for models with (1 | player) etc.)
# ============================================================

  set_prior("normal(-4.6, 0.3)", class = "Intercept"),
  set_prior("student_t(3, 0, 1)", class = "sd", group = "player",  lb = 0),
  set_prior("student_t(3, 0, 1)", class = "sd", group = "gamenu",  lb = 0),
  set_prior("student_t(3, 0, 1)", class = "sd", group = "section", lb = 0)
)

priors_col1 <- c(
  priors_base,
  set_prior("normal(0, 0.08)", class = "b", coef = "plaavspeed_c")
)

priors_col2 <- c(
  priors_base,
  set_prior("normal(0, 0.08)",  class = "b", coef = "plaavspeed_c"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.06)",  class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c")
)

priors_col3 <- c(
  priors_base,
  set_prior("normal(0, 0.08)",  class = "b", coef = "plaavspeed_c"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.06)",  class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c"),
  set_prior("normal(0, 0.15)",  class = "b", coef = "scoregap_pre_c"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteamgoal_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dplagoal_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteamred_pre"),
  set_prior("normal(0, 0.4)",   class = "b", coef = "dteaminj_sec_pre"),
  set_prior("normal(0, 0.025)", class = "b", coef = "team_avespeed_pre_c")
)

priors_col4 <- c(
  priors_col3,
  set_prior("normal(0, 4.8e-05)", class = "b", coef = "team_crowd_c"),
  set_prior("normal(0, 4.8e-05)", class = "b", coef = "opp_crowd_c"),
  set_prior("normal(0, 1.5)",     class = "b", coef = "p_odds_team_c"),
  set_prior("normal(0, 2.5)",     class = "b", coef = "p_odds_draw_c")
)

#ab hier muss noch bearbeitet werden 
priors_col5 <- c(
  priors_base,
  set_prior("normal(0, 0.10)", class = "b", coef = "playersprints_c"),
  set_prior("normal(0, 0.50)", class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.10)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.05)", class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.01)", class = "b", coef = "player_cumsections_2_c"),
  set_prior("normal(0, 0.10)", class = "b", coef = "scoregap_pre_c"),
  set_prior("normal(0, 0.50)", class = "b", coef = "dteamgoal_pre"),
  set_prior("normal(0, 0.50)", class = "b", coef = "dplagoal_pre"),
  set_prior("normal(0, 0.50)", class = "b", coef = "dteamred_pre"),
  set_prior("normal(0, 0.20)", class = "b", coef = "dteaminj_sec_pre"),
  set_prior("normal(0, 0.15)", class = "b", coef = "team_avespeed_pre_c")
)

priors_col6 <- c(
  priors_col5,
  set_prior("normal(0, 0.20)", class = "b", coef = "plaavspeed_c")
)

# Base priors (Table 2, game-level: column 7)
priors_base_col7 <- c(
  set_prior("normal(0, 1.5)", class = "Intercept"),
  set_prior("student_t(3, 0, 1)", class = "sd", group = "player", lb = 0),
  set_prior("student_t(3, 0, 1)", class = "sd", group = "gamenu", lb = 0)
)

priors_col7 <- c(
  priors_base_col7,
  set_prior("normal(0, 0.20)", class = "b", coef = "plaavspeed_game_c"),
  set_prior("normal(0, 0.50)", class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.10)", class = "b", coef = "rank_gap_c")
)


# ============================================================
# (B) FIXED-EFFECTS priors (for models with factor(player) etc.)
# ============================================================

priors_fixef_base <- c(
  set_prior("normal(-4.6, 0.3)", class = "Intercept"),
  set_prior("normal(0, 0.4)",    class = "b")  # applies to all FE dummies unless overridden
) 

### Column 1: dplagoal ~ plaavspeed_c + factor(player) + factor(gamenu) + factor(section)
priors_fixef_col1 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.08)", class = "b", coef = "plaavspeed_c")
)
# We are saying that running at top speed 12km/h changes the probability to score compared to 0km/h between (0.0014 ; 0.064) so 0.1% to 6.4%.  


### Column 2: add dhomegame, rank_gap_c, player_cumsections_c, player_cumsections_2_c
priors_fixef_col2 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.08)", class = "b", coef = "plaavspeed_c"),
# We are saying that running at top speed 12km/h changes the probability to score compared to 0km/h between (0.0014 ; 0.064) so 0.1% to 6.4%. 
  set_prior("normal(0, 0.4)", class = "b", coef = "dhomegame"),
# We are saying that playing at home changes the probability to score to between (0.0045 ; 0.0218) so 0.45% to 2.18%.
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
# We are saying that playing someone that is 13 positions under you changes the probability to score compared to playing someone 13 position above you between (0.0009 ; 0.094) so 0.0% to 9.4%.
  set_prior("normal(0, 0.06)", class = "b", coef = "player_cumsections_c"),
#We are saying that playing 19 sections cumulatively changes the probability to score compared to playing only one section between (0.0009 ; 0.088) so 0.01% to 8.8%.
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c")
#We are saying that playing 361 squaredsection-units cumulatively changes the probability to score compared to playing only one squaredsection-unit between (0.0011 ; 0.079) so 0.01% to 7.9%.
)


### Column 3: add pre-match / lagged controls
### dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c + player_cumsections_c + player_cumsections_2_c +
###             scoregap_pre_c + dteamgoal_pre + dplagoal_pre + dteamred_pre + dteaminj_sec_pre +
###             team_avespeed_pre_c + factors
priors_fixef_col3 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.08)", class = "b", coef = "plaavspeed_c"),
# We are saying that running at top speed 12km/h changes the probability to score compared to 0km/h between (0.0014 ; 0.064) so 0.1% to 6.4%.
set_prior("normal(0, 0.4)", class = "b", coef = "dhomegame"),
# We are saying that playing at home changes the probability to score to between (0.0045 ; 0.0218) so 0.45% to 2.18%.
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
# We are saying that playing someone that is 13 positions under you changes the probability to score compared to playing someone 13 position above you between (0.0009 ; 0.094) so 0.0% to 9.4%.
  set_prior("normal(0, 0.06)", class = "b", coef = "player_cumsections_c"),
#We are saying that playing 19 sections cumulatively changes the probability to score compared to playing only one section between (0.0009 ; 0.088) so 0.01% to 8.8%.
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c"),
#We are saying that playing 361 squaredsection-units cumulatively changes the probability to score compared to playing only one squaredsection-unit between (0.0011 ; 0.079) so 0.01% to 7.9%.
  set_prior("normal(0, 0.15)", class = "b", coef = "scoregap_pre_c"),
#We are saying that having a score advantage of 4 goals changes the probability to score compared having 4 less goals than the oppononte to between (0.009 ; 0.098) so 0.01% to 9.8%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dteamgoal_pre"),
#We are saying that the team scoring a goal in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dplagoal_pre"),
#We are saying that scoring a goal yourself in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dteamred_pre"),
#We are saying that your team getting a red card in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dteaminj_sec_pre"),
#We are saying that your team getting a injury in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.025)", class = "b", coef = "team_avespeed_pre_c")
#We are saying that a team running 49km/h in the section before changes the likelyhood of scoring agoal compared to team running 0km/h between (0.0008 ; 0.103) so 0.01% to 10,3%.
)



### Column 4: plus crowd & betting odds
### dplagoal ~ plaavspeed_c + ... + team_crowd_c + opp_crowd_c + p_odds_team_c + p_odds_draw_c + factors
priors_fixef_col4 <- c(
  priors_fixef_col3,  # all priors from col3 +
  set_prior("normal(0, 4.8e-05)", class = "b", coef = "team_crowd_c"),
#We are saying that a team having  25180 home crowd members changes the probability of scoring a goal compared to having 6 homecrowd members by (0.0008 ; 0.10) so 0.01% to 10%.
  set_prior("normal(0, 4.8e-05)", class = "b", coef = "opp_crowd_c"),
#We are saying that a team having  25180 opp crowd members changes the probability of scoring a goal compared to having 6 oppcrowd members by (0.0008 ; 0.10) so 0.01% to 10%.
  set_prior("normal(0, 1.5)", class = "b", coef = "p_odds_team_c"),
#We are saying that a team having as 0.75 chance of wining compared to 0.05 changes the probability of a goal to be between  (0.00q ; 0.08) so 0.0q% to 8.6%.
  set_prior("normal(0, 2.5)", class = "b", coef = "p_odds_draw_c")
#We are saying that a team having as 0.33 chance of an draw compared to 0.14changes the probability of a goal to be between  (0.003 ; 0.025) so 0.03% to 2.5%.
)

priors_fixef_col4 <- c(
  priors_fixef_col3,
  set_prior("normal(0, 0.05)", class = "b", coef = "team_crowd_c"),
  set_prior("normal(0, 0.05)", class = "b", coef = "opp_crowd_c"),
  set_prior("normal(0, 0.30)", class = "b", coef = "p_odds_team_c"),
  set_prior("normal(0, 0.30)", class = "b", coef = "p_odds_draw_c")
)

priors_fixef_col5 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.10)", class = "b", coef = "playersprints_c"),
  set_prior("normal(0, 0.50)", class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.10)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.05)", class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.01)", class = "b", coef = "player_cumsections_2_c"),
  set_prior("normal(0, 0.10)", class = "b", coef = "scoregap_pre_c"),
  set_prior("normal(0, 0.50)", class = "b", coef = "dteamgoal_pre"),
  set_prior("normal(0, 0.50)", class = "b", coef = "dplagoal_pre"),
  set_prior("normal(0, 0.50)", class = "b", coef = "dteamred_pre"),
  set_prior("normal(0, 0.20)", class = "b", coef = "dteaminj_sec_pre"),
  set_prior("normal(0, 0.15)", class = "b", coef = "team_avespeed_pre_c")
)

priors_fixef_col6 <- c(
  priors_fixef_col5,
  set_prior("normal(0, 0.20)", class = "b", coef = "plaavspeed_c")
)

# Column 7 (game-level): no "sd" priors; still include generic b for the FE dummies
priors_fixef_base_col7 <- c(
  set_prior("normal(0, 1.5)", class = "Intercept"),
  set_prior("normal(0, 0.8)", class = "b")
)

priors_fixef_col7 <- c(
  priors_fixef_base_col7,
  set_prior("normal(0, 0.20)", class = "b", coef = "plaavspeed_game_c"),
  set_prior("normal(0, 0.50)", class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.10)", class = "b", coef = "rank_gap_c")
)

```

```{r}
threads_per_chain <- 2
Sys.setenv(STAN_NUM_THREADS = threads_per_chain)
```



# PP Check Fixed effects Table 2 ---- 
```{r}
m_col1_prior <- brm(
  dplagoal ~ plaavspeed_c + factor(player) + factor(gamenu) + factor(section),
  data      = data_col1,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col1,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col1 <- pp_check(m_col1_prior, ndraws = 100)
ggsave("col1_ppcheck.png", p_col1)

saveRDS(m_col1_prior, "m_col1_prior.rds")

```

```{r}
m_col2_prior <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col2,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col2,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col2 <- pp_check(m_col2_prior, ndraws = 100)
ggsave("col2_ppcheck.png", p_col2)

saveRDS(m_col2_prior, "m_col2_prior.rds")
```

```{r}
m_col3_prior <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col3,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col3 <- pp_check(m_col3_prior, ndraws = 100)
ggsave("col3_ppcheck.png", p_col3)

saveRDS(m_col3_prior, "m_col3_prior.rds")

```


```{r}
m_col4_prior <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    team_crowd_c + opp_crowd_c + p_odds_team_c + p_odds_draw_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col4,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col4,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col4 <- pp_check(m_col4_prior, ndraws = 100)
ggsave("col4_ppcheck.png", p_col4)

saveRDS(m_col4_prior, "m_col4_prior.rds")
```


# Fixed Effects Table 2 ----
```{r}
# ---- FIXED EFFECTS (Table 2), cmdstanr backend, save model + LOO/WAIC ----

# ---------- Column 1 ----------
m_col1 <- brm(
  dplagoal ~ plaavspeed_c + factor(player) + factor(gamenu) + factor(section),
  data      = data_col1,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col1,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col1 <- add_criterion(m_col1, c("loo", "waic"))
saveRDS(m_col1, "m_col1.rds")

# ---------- Column 2 ----------
m_col2 <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col2,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col2,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col2 <- add_criterion(m_col2, c("loo", "waic"))
saveRDS(m_col2, "m_col2.rds")

# ---------- Column 3 ----------
m_col3 <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col3,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col3 <- add_criterion(m_col3, c("loo", "waic"))
saveRDS(m_col3, "m_col3.rds")

# ---------- Column 4 ----------
m_col4 <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    team_crowd_c + opp_crowd_c + p_odds_team_c + p_odds_draw_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col4,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col4,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col4 <- add_criterion(m_col4, c("loo", "waic"))
saveRDS(m_col4, "m_col4.rds")
```

```{r}

# ---------- Column 5 ----------
m_col5 <- brm(
  dplagoal ~ playersprints_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col5,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col5 <- add_criterion(m_col5, c("loo", "waic"))
saveRDS(m_col5, "m_col5.rds")

# ---------- Column 6 ----------
m_col6 <- brm(
  dplagoal ~ plaavspeed_c + playersprints_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col6,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col6 <- add_criterion(m_col6, c("loo", "waic"))
saveRDS(m_col6, "m_col6.rds")

# ---------- Column 7 ----------
m_col7 <- brm(
  goal_game ~ plaavspeed_game_c + dhomegame + rank_gap_c +
    factor(player) + factor(gamenu),
  data      = data_col7,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col7,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col7 <- add_criterion(m_col7, c("loo", "waic"))
saveRDS(m_col7, "m_col7.rds")

```



# PP Check Hierarchical effects Table 2 ----
```{r}
m_col1_hier_prior <- brm(
  dplagoal ~ plaavspeed_c + (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col1,
  family    = bernoulli(link = "logit"),
  prior     = priors_col1,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col1_hier <- pp_check(m_col1_hier_prior, ndraws = 100)
ggsave("col1_ppcheck_hier.png", p_col1_hier)

saveRDS(m_col1_hier_prior, "m_col1_hier_prior.rds")

```

```{r}
m_col2_hier_prior <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col2,
  family    = bernoulli(link = "logit"),
  prior     = priors_col2,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col2_hier <- pp_check(m_col2_hier_prior, ndraws = 100)
ggsave("col2_ppcheck_hier.png", p_col2_hier)

saveRDS(m_col2_hier_prior, "m_col2_hier_prior.rds")

```

```{r}
m_col3_hier_prior <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_col3,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col3_hier <- pp_check(m_col3_hier_prior, ndraws = 100)
ggsave("col3_ppcheck_hier.png", p_col3_hier)

saveRDS(m_col3_hier_prior, "m_col3_hier_prior.rds")

```

```{r}
m_col4_hier_prior <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    team_crowd_c + opp_crowd_c + p_odds_team_c + p_odds_draw_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col4,
  family    = bernoulli(link = "logit"),
  prior     = priors_col4,
  sample_prior = "only",
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

p_col4_hier <- pp_check(m_col4_hier_prior, ndraws = 100)
ggsave("col4_ppcheck_hier.png", p_col4_hier)

saveRDS(m_col4_hier_prior, "m_col4_hier_prior.rds")

```

# Hierarchical models Table 2 ----
```{r}
# ---- HIERARCHICAL / MULTILEVEL (Table 2), cmdstanr backend, save model + LOO/WAIC ----

# ---------- Column 1 ----------
m_col1_hier <- brm(
  dplagoal ~ plaavspeed_c + (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col1,
  family    = bernoulli(link = "logit"),
  prior     = priors_col1,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col1_hier <- add_criterion(m_col1_hier, c("loo", "waic"))
saveRDS(m_col1_hier, "m_col1_hier.rds")

# ---------- Column 2 ----------
m_col2_hier <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col2,
  family    = bernoulli(link = "logit"),
  prior     = priors_col2,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col2_hier <- add_criterion(m_col2_hier, c("loo", "waic"))
saveRDS(m_col2_hier, "m_col2_hier.rds")

# ---------- Column 3 ----------
m_col3_hier <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_col3,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col3_hier <- add_criterion(m_col3_hier, c("loo", "waic"))
saveRDS(m_col3_hier, "m_col3_hier.rds")

# ---------- Column 4 ----------
m_col4_hier <- brm(
  dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    team_crowd_c + opp_crowd_c + p_odds_team_c + p_odds_draw_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col4,
  family    = bernoulli(link = "logit"),
  prior     = priors_col4,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col4_hier <- add_criterion(m_col4_hier, c("loo", "waic"))
saveRDS(m_col4_hier, "m_col4_hier.rds")
```

```{r}

# ---------- Column 5 ----------
m_col5_hier <- brm(
  dplagoal ~ playersprints_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_col5,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col5_hier <- add_criterion(m_col5_hier, c("loo", "waic"))
saveRDS(m_col5_hier, "m_col5_hier.rds")

# ---------- Column 6 ----------
m_col6_hier <- brm(
  dplagoal ~ plaavspeed_c + playersprints_c + dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre + team_avespeed_pre_c +
    (1 | player) + (1 | gamenu) + (1 | section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_col6,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col6_hier <- add_criterion(m_col6_hier, c("loo", "waic"))
saveRDS(m_col6_hier, "m_col6_hier.rds")

# ---------- Column 7 ----------
m_col7_hier <- brm(
  goal_game ~ plaavspeed_game_c + dhomegame + rank_gap_c +
    (1 | player) + (1 | gamenu),
  data      = data_col7,
  family    = bernoulli(link = "logit"),
  prior     = priors_col7,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col7_hier <- add_criterion(m_col7_hier, c("loo", "waic"))
saveRDS(m_col7_hier, "m_col7_hier.rds")

```



# Intercept Only Table 2 ----
```{r}
# ---- Intercept-only (Table 2 baseline), cmdstanr backend, save + LOO/WAIC ----
threads_per_chain <- 2
Sys.setenv(STAN_NUM_THREADS = threads_per_chain)

m_intercept <- brm(
  dplagoal ~ 1,
  data      = data_base,
  family    = bernoulli(link = "logit"),
  prior     = set_prior("normal(-4.6, 0.3)", class = "Intercept"),
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)

m_intercept <- add_criterion(m_intercept, c("loo", "waic"))
saveRDS(m_intercept, "m_intercept.rds")

```




# Looser Priors Table 2 ----
```{r}
# ---- PRIOR SENSITIVITY (LOOSER PRIORS) for Table 2: Columns 1 and 6 (Fixed Effects) ----
threads_per_chain <- 2
Sys.setenv(STAN_NUM_THREADS = threads_per_chain)

# looser FE base priors (applies to all factor() dummies + slopes unless overridden)
priors_fixef_base_loose <- c(
  set_prior("normal(-4.6, 0.6)", class = "Intercept"),
  set_prior("normal(0, 0.8)",    class = "b")
)

# ---- Column 1: dplagoal ~ plaavspeed_c + FE ----
priors_fixef_col1_loose <- c(
  priors_fixef_base_loose,
  set_prior("normal(0, 0.8)", class = "b", coef = "plaavspeed_c")
)

m_col1_loose <- brm(
  dplagoal ~ plaavspeed_c + factor(player) + factor(gamenu) + factor(section),
  data      = data_col1,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col1_loose,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col1_loose <- add_criterion(m_col1_loose, c("loo", "waic"))
saveRDS(m_col1_loose, "m_col1_loose.rds")

# ---- Column 6: dplagoal ~ plaavspeed_c + playersprints_c + controls + FE ----
priors_fixef_col6_loose <- c(
  priors_fixef_base_loose,
  set_prior("normal(0, 0.8)", class = "b", coef = "plaavspeed_c"),
  set_prior("normal(0, 0.8)", class = "b", coef = "playersprints_c"),
  set_prior("normal(0, 1.0)", class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.3)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.2)", class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.1)", class = "b", coef = "player_cumsections_2_c"),
  set_prior("normal(0, 0.3)", class = "b", coef = "scoregap_pre_c"),
  set_prior("normal(0, 1.0)", class = "b", coef = "dteamgoal_pre"),
  set_prior("normal(0, 1.0)", class = "b", coef = "dplagoal_pre"),
  set_prior("normal(0, 1.0)", class = "b", coef = "dteamred_pre"),
  set_prior("normal(0, 0.5)", class = "b", coef = "dteaminj_sec_pre"),
  set_prior("normal(0, 0.5)", class = "b", coef = "team_avespeed_pre_c")
)

m_col6_loose <- brm(
  dplagoal ~ plaavspeed_c + playersprints_c +
    dhomegame + rank_gap_c +
    player_cumsections_c + player_cumsections_2_c +
    scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
    dteamred_pre + dteaminj_sec_pre +
    team_avespeed_pre_c +
    factor(player) + factor(gamenu) + factor(section),
  data      = data_col356,
  family    = bernoulli(link = "logit"),
  prior     = priors_fixef_col6_loose,
  backend   = "cmdstanr",
  threads   = threading(threads_per_chain),
  chains    = 4, iter = 2000, warmup = 1000, cores = 4,
  save_pars = save_pars(all = TRUE)
)
m_col6_loose <- add_criterion(m_col6_loose, c("loo", "waic"))
saveRDS(m_col6_loose, "m_col6_loose.rds")

```



# Table 3
##Intercept-only
```{r}
m_t3_intercept <- brm(
  dteamwin ~ 1, 
  data    = data_teamxgame,
  family  = bernoulli(link = "logit"),
  sample_prior = "only",
  prior = prior( normal(0, 1), class = Intercept),      
  iter    = 1000, warmup = 500, chains = 4, cores = 4
)

pp_check(m_t3_intercept)

#saveRDS(m_t3_intercept, "m_t3_intercept.rds")

```
```{r}
readRDS("m_t3_intercept.rds")
mean(data_teamxgame$dteamwin)
```
Intercept estimate: -0.48 --> 38.2%
Mean also ~38%


## Add Team Names for table 3
```{r}
# add team names to the data for table 3
library(readxl)
team_names <- read_excel("TeamID.xlsx")


#teams don't play the same amount of games, there are 26 games in the first round then they are divided in top teams and bottom teams. isolate the 26 first games (Appendix A.2, in the paper) the regular season



# team rank in game 26 and game 27 because rank variable is the rank before the game

team_rank_r26 <- data %>%
  filter(played == 26) %>%           # only round 26
  distinct(team, team_rank)          # one row per team with its rank

team_rank_r27 <- data %>%
  filter(played == 27) %>%           # only round 27
  distinct(team, team_rank)          # one row per team with its rank

# compared with table "Positions by round" from https://en.wikipedia.org/wiki/2017%E2%80%9318_Israeli_Premier_League#Regular_season_table  -> team_rank_r27 is the right one

# add the israeli name to round 27
team_rank_r27 <- team_rank_r27 %>%
  left_join(
    team_names %>%
      select(team_rank, israeli_name = `händisch israeli name`),
    by = "team_rank"
  )

#add the israeli name to main data
data <- data %>%
  left_join(
    team_rank_r27 %>%
      select(team, israeli_name),
    by = "team"
  )

```


```{r}
##### First, aggregate on the team x game level #####

data_teamxgame <- data %>% 
  group_by(gamenu, team, opponent) %>%  # we keep opponent to include pair fixed effects
  summarise(
    teamdist_game = sum(teamdist_game, na.rm = TRUE),
    oppdist_game  = sum(oppdist_game,  na.rm = TRUE),
    dteamwin      = unique(dteamwin),     # same for all players on team
    team_crowd        = unique(team_crowd),
    opp_crowd         = unique(opp_crowd),
    rank_gap          = unique(rank_gap),
    dhomegame         = unique(dhomegame),
    danyteamredc_game = unique(danyteamredc_game),
    israeli_name      = unique(israeli_name),
    played            = unique(played),
    p_odds_team       = unique(p_odds_team),
    p_odds_draw       = unique(p_odds_draw),
    .groups = "drop"
  ) %>% 
  mutate(
    distance_ratio = teamdist_game / oppdist_game,
    total_distance = teamdist_game + oppdist_game,
    # create pair fixed effects before dropping one team per game
    team_low  = pmin(team, opponent),
    team_high = pmax(team, opponent),
    pair      =paste(team_low, team_high, sep = "_")
  ) %>%
  group_by(gamenu) %>%
  slice_min(israeli_name) %>%      # keep only the team which comes first in the hebrew alpahbet
  ungroup()
```

pair fixed effects
```{r}
data_teamxgame <- data_teamxgame %>%
  mutate(
    team_low  = pmin(team, opponent),
    team_high = pmax(team, opponent),
    pair = paste0(team_low, "_", team_high)
  )
#we have 14 teams, 14 over 2 = 91 possible matchups,verify below
length(unique(data_teamxgame$pair))
```


```{r}
# ---------- Create derived variables and center continuous regressors ----------
data_teamxgame <- data_teamxgame %>%
  mutate(
    distance_ratio = ifelse(is.na(distance_ratio), teamdist_game / ifelse(oppdist_game==0, NA, opp_dist),     distance_ratio),
    total_distance   = teamdist_game + oppdist_game,
    # center continuous predictors (mean-centering)
    teamdist_c = teamdist_game - mean(teamdist_game, na.rm = TRUE),
    oppdist_c  = oppdist_game  - mean(oppdist_game,  na.rm = TRUE),
    distance_ratio_c  = distance_ratio - mean(distance_ratio, na.rm = TRUE),
    total_distance_c  = total_distance - mean(total_distance, na.rm = TRUE),
    rank_gap_c        = rank_gap - mean(rank_gap, na.rm = TRUE),
    team_crowd_c      = team_crowd - mean(team_crowd, na.rm = TRUE),
    opp_crowd_c       = opp_crowd - mean(opp_crowd, na.rm = TRUE),
    p_odds_team_c     = p_odds_team - mean(p_odds_team, na.rm = TRUE),
    p_odds_draw_c     = p_odds_draw - mean(p_odds_draw, na.rm = TRUE)
  )

#datasets for each column 
table3_col1 <- data_teamxgame %>%
  drop_na(dteamwin, teamdist_game, oppdist_game)  

table3_col2 <- data_teamxgame %>%
  drop_na(dteamwin, teamdist_game, oppdist_game, dhomegame, rank_gap)

table3_col3 <- data_teamxgame %>%
  drop_na(dteamwin, teamdist_game, oppdist_game, dhomegame, rank_gap, team_crowd_c, opp_crowd_c, p_odds_team_c, p_odds_draw_c)

table3_col4 <- data_teamxgame %>%
  # restrict to games without red cards
  filter(is.na(danyteamredc_game) | danyteamredc_game == 0) %>%
  drop_na(dteamwin, teamdist_game, oppdist_game, dhomegame, rank_gap)

table3_col5 <- data_teamxgame %>%
  drop_na(dteamwin, distance_ratio, dhomegame, rank_gap)

table3_col6 <- data_teamxgame %>%
  drop_na(dteamwin, distance_ratio, total_distance, dhomegame, rank_gap)

# double-check sample sizes
sapply(list(table3_col1, table3_col2, table3_col3, table3_col4, table3_col5, table3_col6), nrow)
```

slightly different sample sizes compared to the ones from table 3 in the paper (why?)

#col1
```{r}
#set weakly informative priors that fit logit scale 
priors_t3c1 <- c( 
  set_prior("normal(-0.48, 1.23)", class = "Intercept"), # 38% baseline chance, 95% of values will be between [0.05, 0.95], this is wider than the BO range of 6% to 77% 
  set_prior("normal(0, 0.0001)", class = "b", coef = "teamdist_c"),
# we are saying that a team that runs 26.649 compared to a team that runs 19.426 will change the probability of winning between (0.12 , 0.72 ) so from 12% to 72%
  set_prior("normal(0, 0.0001)", class = "b", coef = "oppdist_c")
# we are saying that a team whose opponent runs 26.515 compared to a team that runs 20.514 will change the probability of winning between (0.15 , 0.67 ) so from 15% to 67%
)
```


```{r}
m_t3col1_prior <- brm( 
  dteamwin ~ teamdist_c + oppdist_c + (1 | pair) + (1 | played),
  data = table3_col1, 
  family = bernoulli(link = "logit"),
  sample_prior = "only",
  prior = priors_t3c1,
  iter = 1000, warmup = 500, chains = 4, cores = 4 
)
```

```{r}
#prior checks
pp_check(m_t3col1_prior, ndraws = 100)          
yrep_small <- posterior_predict(m_t3col1_prior, ndraws = 100)
hist(rowMeans(yrep_small))
plot(m_t3col1_prior, N = 1000)                             
```

```{r}
m_t3col1 <- brm(
  dteamwin ~ teamdist_c + oppdist_c + (1 | pair) + (1 | played),
  data    = table3_col1,
  family  = bernoulli(link = "logit"),
  prior   = priors_t3,
  iter    = 4000, warmup = 2000, chains = 4, cores = 4
)
```

```{r}
summary(m_t3col1)
pp_check(m_t3col1, ndraws = 100)
#conditional_effects(m_t3col1, spaghetti = TRUE, ndraws = 200)
plot(m_t3col1)
```

```{r}
cor(table3_col1 %>% select(teamdist_c, oppdist_c, distance_ratio_c, total_distance_c), use = "pairwise.complete.obs")
```
is multicollinearity causing problems here in col1?



#col2
```{r}
priors_t3c2 <- c( 
  set_prior("normal(-0.48, 1.23)", class = "Intercept"), # 38% baseline chance, 95% of values will be between [0.05, 0.95], this is wider than the BO range of 6% to 77% 
  set_prior("normal(0, 0.0001)", class = "b", coef = "teamdist_c"),
# we are saying that a team that runs 26.649 compared to a team that runs 19.426 will change the probability of winning between (0.12 , 0.72 ) so from 12% to 72%
  set_prior("normal(0, 0.0001)", class = "b", coef = "oppdist_c"),
# we are saying that a team whose opponent runs 26.515 compared to a team that runs 20.514 will change the probability of winning between (0.15 , 0.67 ) so from 15% to 67%
  set_prior("normal(0, 0.5)",  class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.3)", class = "b", coef = "rank_gap_c")
)
```

```{r}
m_t3col2 <- brm(
  dteamwin ~ teamdist_c + oppdist_c + (1 | pair) + (1 | played) +
    dhomegame +
    rank_gap_c,     
  data    = table3_col2,
  family  = bernoulli(link = "logit"),
  prior   = priors_t3c2,
  iter    = 4000, warmup = 2000, chains = 4, cores = 4,
  control = list(max_treedepth = 15, adapt_delta = 0.95) #increase treedepth to avoid problems during warmup
)
summary(m_t3col2)
#plot(m_t3col2)
```


#col3
```{r}
priors_t3c3 <- c( 
  set_prior("normal(0, 1)", class = "Intercept"),
  set_prior("normal(0, 1)", class = "b", coef = "teamdist_c"),
  set_prior("normal(0, 1)", class = "b", coef = "oppdist_c"),
  set_prior("normal(0, 0.5)",  class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.3)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.3)", class = "b", coef = "team_crowd_c"),
  set_prior("normal(0, 0.3)", class = "b", coef = "opp_crowd_c"),
  set_prior("normal(0, 0.5)", class = "b", coef = "p_odds_team_c"),
  set_prior("normal(0, 0.5)", class = "b", coef = "p_odds_draw_c")
)
```

```{r}
m_t3col3 <- brm(
  dteamwin ~ 
    teamdist_c + 
    oppdist_c + 
    team_crowd_c +
    opp_crowd_c +
    p_odds_team_c +
    p_odds_draw_c +
    (1 | pair) + 
    (1 | played) +
    dhomegame +
    rank_gap_c,     
  data    = table3_col3,
  family  = bernoulli(link = "logit"),
  prior   = priors_t3c3,
  iter    = 4000, warmup = 2000, chains = 4, cores = 4,
  control = list(max_treedepth = 15, adapt_delta = 0.95) #increase treedepth to avoid problems during warmup
)
summary(m_t3col3)
#plot(m_t3col3)
```

#col4
```{r}
priors_t3c4 <- c( 
  set_prior("normal(0, 1)", class = "Intercept"),
  set_prior("normal(0, 1)", class = "b", coef = "teamdist_c"),
  set_prior("normal(0, 1)", class = "b", coef = "oppdist_c"),
  set_prior("normal(0, 0.5)",  class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.3)", class = "b", coef = "rank_gap_c")
)
```

```{r}
m_t3col4 <- brm(
  dteamwin ~ teamdist_c + oppdist_c + (1 | pair) + (1 | played) +
    dhomegame +
    rank_gap_c,     
  data    = table3_col4,
  family  = bernoulli(link = "logit"),
  prior   = priors_t3c4,
  iter    = 4000, warmup = 2000, chains = 4, cores = 4,
  control = list(max_treedepth = 15, adapt_delta = 0.95) #increase treedepth to avoid problems during warmup
)
summary(m_t3col4)
#plot(m_t3col4)
```



#col5

below we impose reasonably skeptical priors. What the priors say:
Distance ratio matters, but not absurdly.
Home advantage is modest.
Rank gap matters a little, but not too much.
If distance_ratio_c large ->data-driven
If it shrinks strongly ->prior-sensitive result
Regression Coefficients:
                 Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
Intercept           -0.77      0.24    -1.26    -0.29 1.00    10422     5941
distance_ratio_c     0.01      0.10    -0.18     0.20 1.00    16599     5784
dhomegame            0.29      0.26    -0.23     0.81 1.00    16276     6145
rank_gap_c          -0.18      0.04    -0.26    -0.11 1.00    10318     5311

--> given reasonable skepticism, the data do not strongly identify a positive causal effect of relative distance
--> When estimating a game-level win probability with pair and round effects, the effect of relative running distance is highly sensitive to modeling assumptions. Under weakly informative priors and partial pooling, the posterior mass concentrates near zero, indicating limited identifying variation once unobserved heterogeneity is accounted for.

If we set our priors such that we allow/expect huge effects (set_prior("normal(0, 3)", class = "b", coef = "distance_ratio_c"), we can produce point estimates for distance_ratio_c similar to those in the paper (and *** significance level). This indicates that the data are weakly informative about the effect of distance_ratio_c and therefore the posterior is very sensitive to the prior. Priors should therefore be set in a way that seems intuitive -> "home advantage exists but is modest, rank gap matters (how much though), distance_ratio matters very little?"
After controlling for pair and round heterogeneity, there is very little variation left which can be uniquely attributed to distance_ratio. (prior heavily dominates the likelihood)

Ideas for why the authors get such a strong estimate for distance_ratio in the paper:
classical fixed-effects logit assumes asymptotic identification, does not penalize overparameterization, reports a single MLE even when likelihood curvature is weak
--> it answers: "what coefficient maximizes the likelihood?" rather than "how confident should we be about this magnitude?"
In our bayesian model, the data provide little to no evidence for a sizable effect of distance_ratio on winning probabilities given the fixed effects. Under reasonable priors, the posterior mass is concentratred near zero.
Important: the model doesnt prove that the effect is weak! It instead says that the data do not support a strong effect of distance_ratio on winning probability.


```{r}
priors_t3c5 <-c(
  set_prior("normal(0, 0.5)",  class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.3)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.1)", class = "b", coef = "distance_ratio_c") # wider priors (normal 0, 2-3) produce results close to paper, otherwise effect is shrunk towards zero
)
```


```{r}
m_t3col5 <- brm(
  dteamwin ~ distance_ratio_c + (1 | pair) + (1 | played) +
    dhomegame +
    rank_gap_c,     
  data    = table3_col5,
  family  = bernoulli(link = "logit"),
  prior   = priors_t3c5,
  iter    = 4000, warmup = 2000, chains = 4, cores = 4,
  control = list(max_treedepth = 15, adapt_delta = 0.95) #increase treedepth to avoid problems during warmup
)
```

```{r}
summary(m_t3col5)
#pp_check(m_t3col5, ndraws = 100)
#conditional_effects(m_t3col1, spaghetti = TRUE, ndraws = 200)
#plot(m_t3col5)
```

#col6
how do we choose priors?
distance_ratio --> relative effort (could have moderate effect)
total_distance --> game intensity/pace (probably weak effect if at all? guessing its near zero)
aim for weak regularization

```{r}
priors_t3c6 <-c(
  set_prior("normal(0, 0.5)",  class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.3)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 2)", class = "b", coef = "distance_ratio_c"),
  set_prior("normal(0, 0.5)", class = "b", coef = "total_distance_c")
)
```

```{r}
m_t3col6 <- brm(
  dteamwin ~ distance_ratio_c + total_distance_c + (1 | pair) + (1 | played) +
    dhomegame +
    rank_gap_c,     
  data    = table3_col6,
  family  = bernoulli(link = "logit"),
  prior   = priors_t3c6,
  iter    = 4000, warmup = 2000, chains = 4, cores = 4,
  control = list(max_treedepth = 15, adapt_delta = 0.95) #increase treedepth to avoid problems during warmup
)
summary(m_t3col6)
#plot(m_t3col6)
```



#loo evaluation for table 3
```{r}
library(loo)
#gibt es einen workaround für die unterschiedlich großen sample sizes? loo braucht gleiche observation counts
# compute LOO objects
loo_t3c1 <- loo(m_t3col1)
loo_t3c2 <- loo(m_t3col2)
loo_t3c3 <- loo(m_t3col3)
loo_t3c4 <- loo(m_t3col4)
loo_t3c5 <- loo(m_t3col5)
loo_t3c6 <- loo(m_t3col6)

# compare models
loo_compare(
  loo_t3c1,
  loo_t3c2,
 # loo_t3c3, 180 obs
 # loo_t3c4, 152 ob
  loo_t3c5,
  loo_t3c6
)

```
what do the loo results tell us?
---> If |Δelpd| < 2 × SE ---> differences are not decisive
We evaluate model performance using approximate leave-one-out cross-validation (LOO). Since Table 3 specifications rely on different subsamples by construction, LOO comparisons are conducted only among models estimated on the same sample (N = 228).

The results indicate that the specification in Column (2), which includes team and opponent distance together with home advantage and rank gap controls, yields the highest expected log predictive density. Models relying solely on distance measures or relative distance perform worse in terms of out-of-sample prediction, while differences between Columns (2) and (6) are not statistically decisive. Overall, distance-based variables add limited predictive value once pair and round heterogeneity are accounted for.



# Ich glaube ab hier wird nicht mehr beötigt ----
## LD Table 2 priors

```{r}
# function to check the p given a chosen coefficient and given intercept alpha 
p_from_beta <- function(alpha, beta, x) {
  exp(alpha + beta * x) / (1 + exp(alpha + beta * x))
}


# function to get the coefficient beta based on desired p given an intercept alpha
beta_from_p <- function(p, alpha, x, eps = 1e-12) {
  # Basic checks
  if (any(is.na(p) | is.na(alpha) | is.na(x))) stop("Inputs contain NA.")
  if (any(p <= 0 | p >= 1)) stop("p must be strictly between 0 and 1.")
  if (any(abs(x) < eps)) stop("x must be non-zero (cannot divide by ~0).")

  # logit(p) = log(p/(1-p))
  logit_p <- log(p / (1 - p))
  (logit_p - alpha) / x
}




## 1) Base priors: used for columns 1–6
priors_fixef_base <- c(
  # same intercept prior as hierarchical cols 1–6
  set_prior("normal(-4.6, 0.3)", class = "Intercept"),
# 1% based on paper
  
  # default for all coefficients (incl. factor dummies), unless overridden below
  set_prior("normal(0, 0.4)",   class = "b")
) 
# We are saying that each dummy changes the probability to score to between (0.0045 ; 0.0218) so 0.45% to 2.18%.  


## 2) Column-specific priors: base + explicit priors for every covariate in that column

### Column 1: dplagoal ~ plaavspeed_c + factor(player) + factor(gamenu) + factor(section)
priors_fixef_col1 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.08)", class = "b", coef = "plaavspeed_c")
)
# We are saying that running at top speed 12km/h changes the probability to score compared to 0km/h between (0.0014 ; 0.064) so 0.1% to 6.4%.  


### Column 2: add dhomegame, rank_gap_c, player_cumsections_c, player_cumsections_2_c
priors_fixef_col2 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.08)", class = "b", coef = "plaavspeed_c"),
# We are saying that running at top speed 12km/h changes the probability to score compared to 0km/h between (0.0014 ; 0.064) so 0.1% to 6.4%. 
  set_prior("normal(0, 0.4)", class = "b", coef = "dhomegame"),
# We are saying that playing at home changes the probability to score to between (0.0045 ; 0.0218) so 0.45% to 2.18%.
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
# We are saying that playing someone that is 13 positions under you changes the probability to score compared to playing someone 13 position above you between (0.0009 ; 0.094) so 0.0% to 9.4%.
  set_prior("normal(0, 0.06)", class = "b", coef = "player_cumsections_c"),
#We are saying that playing 19 sections cumulatively changes the probability to score compared to playing only one section between (0.0009 ; 0.088) so 0.01% to 8.8%.
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c")
#We are saying that playing 361 squaredsection-units cumulatively changes the probability to score compared to playing only one squaredsection-unit between (0.0011 ; 0.079) so 0.01% to 7.9%.
)


### Column 3: add pre-match / lagged controls
### dplagoal ~ plaavspeed_c + dhomegame + rank_gap_c + player_cumsections_c + player_cumsections_2_c +
###             scoregap_pre_c + dteamgoal_pre + dplagoal_pre + dteamred_pre + dteaminj_sec_pre +
###             team_avespeed_pre_c + factors
priors_fixef_col3 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.08)", class = "b", coef = "plaavspeed_c"),
# We are saying that running at top speed 12km/h changes the probability to score compared to 0km/h between (0.0014 ; 0.064) so 0.1% to 6.4%.
set_prior("normal(0, 0.4)", class = "b", coef = "dhomegame"),
# We are saying that playing at home changes the probability to score to between (0.0045 ; 0.0218) so 0.45% to 2.18%.
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
# We are saying that playing someone that is 13 positions under you changes the probability to score compared to playing someone 13 position above you between (0.0009 ; 0.094) so 0.0% to 9.4%.
  set_prior("normal(0, 0.06)", class = "b", coef = "player_cumsections_c"),
#We are saying that playing 19 sections cumulatively changes the probability to score compared to playing only one section between (0.0009 ; 0.088) so 0.01% to 8.8%.
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c"),
#We are saying that playing 361 squaredsection-units cumulatively changes the probability to score compared to playing only one squaredsection-unit between (0.0011 ; 0.079) so 0.01% to 7.9%.
  set_prior("normal(0, 0.15)", class = "b", coef = "scoregap_pre_c"),
#We are saying that having a score advantage of 4 goals changes the probability to score compared having 4 less goals than the oppononte to between (0.009 ; 0.098) so 0.01% to 9.8%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dteamgoal_pre"),
#We are saying that the team scoring a goal in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dplagoal_pre"),
#We are saying that scoring a goal yourself in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dteamred_pre"),
#We are saying that your team getting a red card in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dteaminj_sec_pre"),
#We are saying that your team getting a injury in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.025)", class = "b", coef = "team_avespeed_pre_c")
#We are saying that a team running 49km/h in the section before changes the likelyhood of scoring agoal compared to team running 0km/h between (0.0008 ; 0.103) so 0.01% to 10,3%.
)



### Column 4: plus crowd & betting odds
### dplagoal ~ plaavspeed_c + ... + team_crowd_c + opp_crowd_c + p_odds_team_c + p_odds_draw_c + factors
priors_fixef_col4 <- c(
  priors_fixef_col3,  # all priors from col3 +
  set_prior("normal(0, 4.8e-05)", class = "b", coef = "team_crowd_c"),
#We are saying that a team having  25180 home crowd members changes the probability of scoring a goal compared to having 6 homecrowd members by (0.0008 ; 0.10) so 0.01% to 10%.
  set_prior("normal(0, 4.8e-05)", class = "b", coef = "opp_crowd_c"),
#We are saying that a team having  25180 opp crowd members changes the probability of scoring a goal compared to having 6 oppcrowd members by (0.0008 ; 0.10) so 0.01% to 10%.
  set_prior("normal(0, 1.5)", class = "b", coef = "p_odds_team_c"),
#We are saying that a team having as 0.75 chance of wining compared to 0.05 changes the probability of a goal to be between  (0.00q ; 0.08) so 0.0q% to 8.6%.
  set_prior("normal(0, 2.5)", class = "b", coef = "p_odds_draw_c")
#We are saying that a team having as 0.33 chance of an draw compared to 0.14changes the probability of a goal to be between  (0.003 ; 0.025) so 0.03% to 2.5%.
)

### Column 5: sprint effort instead of plaavspeed_c
### dplagoal ~ playersprints_c + same controls as col3 + factors
priors_fixef_col5 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.1)", class = "b", coef = "playersprints_c"),
#we are saying that a player having ran 12 times in this section compared to zero changes the probability of scoring between (0.0009 ; 0.9975) so 0.1% to 10%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dhomegame"),
# We are saying that playing at home changes the probability to score to between (0.0045 ; 0.0218) so 0.45% to 2.18%.
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
# We are saying that playing someone that is 13 positions under you changes the probability to score compared to playing someone 13 position above you between (0.0009 ; 0.094) so 0.0% to 9.4%.
  set_prior("normal(0, 0.06)", class = "b", coef = "player_cumsections_c"),
#We are saying that playing 19 sections cumulatively changes the probability to score compared to playing only one section between (0.0009 ; 0.088) so 0.01% to 8.8%.
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c"),
#We are saying that playing 361 squaredsection-units cumulatively changes the probability to score compared to playing only one squaredsection-unit between (0.0011 ; 0.079) so 0.01% to 7.9%.
  set_prior("normal(0, 0.15)", class = "b", coef = "scoregap_pre_c"),
#We are saying that having a score advantage of 4 goals changes the probability to score compared having 4 less goals than the oppononte to between (0.009 ; 0.098) so 0.01% to 9.8%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dteamgoal_pre"),
#We are saying that the team scoring a goal in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dplagoal_pre"),
#We are saying that scoring a goal yourself in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dteamred_pre"),
#We are saying that your team getting a red card in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dteaminj_sec_pre"),
#We are saying that your team getting a injury in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.025)", class = "b", coef = "team_avespeed_pre_c")
#We are saying that a team running 49km/h in the section before changes the likelyhood of scoring agoal compared to team running 0km/h between (0.0008 ; 0.103) so 0.01% to 10,3%.
)

### Column 6: both plaavspeed_c and playersprints_c + same controls as col5
priors_fixef_col6 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.08)", class = "b", coef = "plaavspeed_c"),
# We are saying that running at top speed 12km/h changes the probability to score compared to 0km/h between (0.0014 ; 0.064) so 0.1% to 6.4%.
  set_prior("normal(0, 0.1)", class = "b", coef = "playersprints_c"),
#we are saying that a player having ran 12 times in this section compared to zero changes the probability of scoring between (0.0009 ; 0.9975) so 0.1% to 10%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dhomegame"),
# We are saying that playing at home changes the probability to score to between (0.0045 ; 0.0218) so 0.45% to 2.18%.
  set_prior("normal(0, 0.045)", class = "b", coef = "rank_gap_c"),
# We are saying that playing someone that is 13 positions under you changes the probability to score compared to playing someone 13 position above you between (0.0009 ; 0.094) so 0.0% to 9.4%.
  set_prior("normal(0, 0.06)", class = "b", coef = "player_cumsections_c"),
#We are saying that playing 19 sections cumulatively changes the probability to score compared to playing only one section between (0.0009 ; 0.088) so 0.01% to 8.8%.
  set_prior("normal(0, 0.003)", class = "b", coef = "player_cumsections_2_c"),
#We are saying that playing 361 squaredsection-units cumulatively changes the probability to score compared to playing only one squaredsection-unit between (0.0011 ; 0.079) so 0.01% to 7.9%.
  set_prior("normal(0, 0.15)", class = "b", coef = "scoregap_pre_c"),
#We are saying that having a score advantage of 4 goals changes the probability to score compared having 4 less goals than the oppononte to between (0.009 ; 0.098) so 0.01% to 9.8%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dteamgoal_pre"),
#We are saying that the team scoring a goal in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dplagoal_pre"),
#We are saying that scoring a goal yourself in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dteamred_pre"),
#We are saying that your team getting a red card in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.4)", class = "b", coef = "dteaminj_sec_pre"),
#We are saying that your team getting a injury in the section before changes the probability to score a goal to be between  0.45% and 2.18%.
  set_prior("normal(0, 0.025)", class = "b", coef = "team_avespeed_pre_c")
#We are saying that a team running 49km/h in the section before changes the likelyhood of scoring agoal compared to team running 0km/h between (0.0008 ; 0.103) so 0.01% to 10,3%.
)

## 3) Column 7: game-level model, wider priors like hierarchical priors_base_col7

#estimate average goal probability in game to get an intercept estimate
mean(data_col7$goal_game) # 11%

priors_fixef_base_col7 <- c(
  set_prior("normal(-2.2, 0.3)", class = "Intercept"),  
# base value based on avg goals per game of around 11%, less or equal variation at game level
  set_prior("normal(0, 0.5)", class = "b")           
# all coefficients by default. We are saying the average factor turning from 0 to 1 changes the probability from ( 0.039 ; 0.231) , so 4% to 23%
)

### Column 7: goal_game ~ plaavspeed_game_c + dhomegame + rank_gap_c + factor(player) + factor(gamenu)
priors_fixef_col7 <- c(
  priors_fixef_base_col7,
  set_prior("normal(0, 0.05)", class = "b", coef = "plaavspeed_game_c"),
# We are saying that increasing the average speed in the game from 2.6 km/h to 10.2km/h changes the scoring probability from (0.04742 , 0.1978) so from 4.7% to 20%
  set_prior("normal(0, 0.5)", class = "b", coef = "dhomegame"),
# We are saying playing at home vs not playing playing at home changes the probability from ( 0.039 ; 0.231) , so 4% to 23%
  set_prior("normal(0, 0.015)", class = "b", coef = "rank_gap_c")
# We are saying playing with 13 ranks higher compared to playing with 13 ranks lower than the opponent changes the probability of winning from (0.0483 , 0.1946) so from 4.8% to 19.4%.
)

```



## Table 2
##Intercept-only model
```{r} 
#needed?
m_intercept <- brm(
  dplagoal ~ 1, 
  data    = data_base,
  family  = bernoulli(link = "logit"),
  #sample_prior = "only",
  prior = prior( normal(0, 3.5), class = Intercept),      
  iter    = 2000, warmup = 1000, chains = 4, cores = 4
)

pp_check(m_intercept)
summary(m_intercept)


posterior_samples( m_intercept ) %>% 
  transmute( alpha = inv_logit_scaled( b_Intercept )) %>% 
  mean_qi()
```

##Preparation
```{r}
#needed?
data_col1s <- data_base %>%
  drop_na(dplagoal, plaavspeed) %>%
  mutate(
    plaavspeed_s = scale(plaavspeed)
  )
```

```{r}
#needed?
priors_fixef_base <- c(
  set_prior("normal(-4.6, 0.3)", class = "Intercept"),
  set_prior("normal(0, 0.4)",   class = "b")
)

priors_fixef_col1 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.4)", class = "b", coef = "plaavspeed_c")
)
```

```{r}
#needed?
sd(data_col1s$plaavspeed)
```

```{r}
#needed?
#set threads for multithreading
Sys.setenv(STAN_NUM_THREADS = threads_per_chain)


m_col1_s <- brm(
  dplagoal ~ plaavspeed_s +
    factor(player) +      
    factor(gamenu) +     
    factor(section), 
  data    = data_col1s,
  family  = bernoulli(link = "logit"),
  #sample_prior = "only",
  prior   = c( prior(normal(-4.6, 0.3), class = "Intercept"),
               prior(normal(0, 0.4), class = "b"),
               prior(normal(0, 0.4681233), class = "b", coef = "plaavspeed_s")),
  iter    = 1000, warmup = 500, chains = 4, cores = 4,
  backend = "cmdstanr",
    threads = threading(threads = threads_per_chain)
)
saveRDS(m_col1_s, "m_col1_s_prior.rds")
```

```{r}
#needed?
col1s <- readRDS("m_col1_s_prior.rds")
col1c <- readRDS("m_col1.rds")

fixef(col1s) - fixef(col1c)

```
