---
title: "Speed Setup"
author: "Lourenco Debonnet"
date: "2025-12-14"
output: html_document
---

## Speed Setups

### install cmdstanr backend 

CmdStanR executes the same Stan model with lower overhead(CmdStan avoids R’s interpreter and memory system during sampling, eliminating R↔C++ boundary crossings and allowing a tight, uninterrupted C++ execution loop.) and supports within-chain parallelization, allowing each MCMC chain to use multiple CPU threads.

```{r}
# 0) optional: only install cmdstanr if missing
if (!requireNamespace("cmdstanr", quietly = TRUE)) {
  install.packages(
    "cmdstanr",
    repos = c("https://mc-stan.org/r-packages/", "https://cloud.r-project.org")
  )
}

library(cmdstanr)

# 1) fix / configure the C++ toolchain (Rtools on Windows)
cmdstanr::check_cmdstan_toolchain(fix = TRUE, quiet = FALSE)

# 2) verify it worked
cmdstanr::check_cmdstan_toolchain()

# 3) install CmdStan itself (one-time, can take a while)
cmdstanr::install_cmdstan()

# 4) verify CmdStan is available
cmdstanr::cmdstan_version()

```


### Check how many processors for parallelization the computer has available

```{r cpu_info_vars, message=FALSE, warning=FALSE}
library(parallel)

# --- Detect CPU resources visible to R ---
max_logical  <- parallel::detectCores(logical = TRUE)
max_physical <- parallel::detectCores(logical = FALSE)

cat("Logical processors available to R:", max_logical, "\n")
cat("Physical cores available to R:   ", max_physical, "\n\n")

# Windows-only: extra confirmation (optional)
if (.Platform$OS.type == "windows") {
  cpu_raw <- try(system("wmic cpu get NumberOfCores,NumberOfLogicalProcessors /value", intern = TRUE), silent = TRUE)
  if (!inherits(cpu_raw, "try-error")) {
    cpu_raw <- cpu_raw[nzchar(trimws(cpu_raw))]
    cat("WMIC (Windows) reported:\n", paste(cpu_raw, collapse = "\n"), "\n\n")
  }
}

# --- Choose brms settings (you can change the caps if you want) ---
# Keep chains at 4 by default for diagnostics, but never exceed logical CPUs
brms_chains <- min(4, max_logical)

# Run chains in parallel up to physical cores (or fewer if chains < cores)
brms_cores  <- min(max_physical, brms_chains)

# For cmdstanr within-chain threading: use remaining logical threads per chain
threads_per_chain <- max(1, floor(max_logical / brms_chains))

cat("Suggested brms settings:\n")
cat("- brms_chains =", brms_chains, "\n")
cat("- brms_cores  =", brms_cores,  " (parallel chains)\n")
cat("- threads_per_chain =", threads_per_chain, " (cmdstanr only)\n")
```

-> jede chain sollte einen core nutzen, wenn man mehr logische einheiten hat, kann man noch zusätzlich multi-threading benuzten, aber cores * multithreading muss kleiner gleich der Anzahl an logischen Einheiten sein



Die Funktionen muss man leicht anpassen
```{r}

library(cmdstanr)   # Für das neue backend


#set threads for multithreading
Sys.setenv(STAN_NUM_THREADS = threads_per_chain)   # um multithreading zu nutzen



m_col1_prior_cmdstan <- brm(
  dplagoal ~ plaavspeed_c +
  factor(player) +      
  factor(gamenu) +     
  factor(section),
  data    = data_col1,
  family  = bernoulli(link = "logit"),
  #sample_prior = "only",
  prior   = priors_fixef_col1,
  iter    = 1000, warmup = 500,
  chains  = brms_chains, cores = brms_cores,    # hier kann man variablen nutzen oder wie im text oben 
  backend = "cmdstanr",                        # das muss dazukommen, um das neue backend zu nutzen
  threads = threading(threads = threads_per_chain)  # das ist für multithreading
)

```

