---
title: "Speed test"
author: "Lourenco Debonnet"
date: "2025-12-13"
output: html_document
---

```{r}
library(tidyverse)
library(haven)
library(ggplot2)
library(dplyr)
library(plotly)
library(tidyr)
library(xtable)
library(knitr)
library(kableExtra)
library(tidybayes)
library(brms)
library(bayesplot)
library(posterior)
```


```{r}
data <- read_dta("original&computed_DB_forrep.dta")
```

```{r}
data_base <- data %>%
  filter(
    dfullsection == 1,          # full sections only
    position %in% c(3, 4),      # midfielders + strikers
    !(section %in% c(10, 20)) 
  )
data_base <- data_base  %>% mutate(player = factor(player),
                                    gamenu = factor(gamenu),
                                    section = factor(section))
```

```{r}
data_col1 <- data_base %>%
  drop_na(dplagoal, plaavspeed) %>%
  mutate(
    plaavspeed_c = plaavspeed - mean(plaavspeed, na.rm = TRUE)
  )
```


# Priors
```{r}
## Fixed-effects priors (for models with factor(player), factor(gamenu), factor(section))
## Designed to be as close as possible in scale to the hierarchical setup.

## 1) Base priors: used for columns 1–6
priors_fixef_base <- c(
  # same intercept prior as hierarchical cols 1–6
  set_prior("normal(-4.6, 0.3)", class = "Intercept"),
  # default for all coefficients (incl. factor dummies), unless overridden below
  set_prior("normal(0, 0.4)",   class = "b")
)

## 2) Column-specific priors: base + explicit priors for every covariate in that column

### Column 1: dplagoal ~ plaavspeed_c + factor(player) + factor(gamenu) + factor(section)
priors_fixef_col1 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.4)", class = "b", coef = "plaavspeed_c")
)


```




# Speed Test

## Speed Setups

### install cmdstanr backend 

CmdStanR executes the same Stan model with lower overhead(CmdStan avoids R’s interpreter and memory system during sampling, eliminating R↔C++ boundary crossings and allowing a tight, uninterrupted C++ execution loop.) and supports within-chain parallelization, allowing each MCMC chain to use multiple CPU threads.

```{r}
# 0) optional: only install cmdstanr if missing
if (!requireNamespace("cmdstanr", quietly = TRUE)) {
  install.packages(
    "cmdstanr",
    repos = c("https://mc-stan.org/r-packages/", "https://cloud.r-project.org")
  )
}

library(cmdstanr)

# 1) fix / configure the C++ toolchain (Rtools on Windows)
cmdstanr::check_cmdstan_toolchain(fix = TRUE, quiet = FALSE)

# 2) verify it worked
cmdstanr::check_cmdstan_toolchain()

# 3) install CmdStan itself (one-time, can take a while)
cmdstanr::install_cmdstan()

# 4) verify CmdStan is available
cmdstanr::cmdstan_version()

```


### Check how many processors for parallelization the computer has available

```{r cpu_info_vars, message=FALSE, warning=FALSE}
library(parallel)

# --- Detect CPU resources visible to R ---
max_logical  <- parallel::detectCores(logical = TRUE)
max_physical <- parallel::detectCores(logical = FALSE)

cat("Logical processors available to R:", max_logical, "\n")
cat("Physical cores available to R:   ", max_physical, "\n\n")

# Windows-only: extra confirmation (optional)
if (.Platform$OS.type == "windows") {
  cpu_raw <- try(system("wmic cpu get NumberOfCores,NumberOfLogicalProcessors /value", intern = TRUE), silent = TRUE)
  if (!inherits(cpu_raw, "try-error")) {
    cpu_raw <- cpu_raw[nzchar(trimws(cpu_raw))]
    cat("WMIC (Windows) reported:\n", paste(cpu_raw, collapse = "\n"), "\n\n")
  }
}

# --- Choose brms settings (you can change the caps if you want) ---
# Keep chains at 4 by default for diagnostics, but never exceed logical CPUs
brms_chains <- min(4, max_logical)

# Run chains in parallel up to physical cores (or fewer if chains < cores)
brms_cores  <- min(max_physical, brms_chains)

# For cmdstanr within-chain threading: use remaining logical threads per chain
threads_per_chain <- max(1, floor(max_logical / brms_chains))

cat("Suggested brms settings:\n")
cat("- brms_chains =", brms_chains, "\n")
cat("- brms_cores  =", brms_cores,  " (parallel chains)\n")
cat("- threads_per_chain =", threads_per_chain, " (cmdstanr only)\n")
```



## normal rstan model, timing
```{r }
set.seed(123)

time_rstan <- system.time({
  m_col1_prior_rstan <- brm(
    dplagoal ~ plaavspeed_c +
    factor(player) +      
    factor(gamenu) +     
    factor(section),
    data    = data_col1,
    family  = bernoulli(link = "logit"),
    #sample_prior = "only",
    prior   = priors_fixef_col1,
    iter    = 1000, warmup = 500,
    chains  = brms_chains, cores = brms_cores,
    backend = "rstan",
    refresh = 0
  )
})

time_rstan
```



## cmdstanr model

```{r}

library(cmdstanr)
set.seed(123)

# no mulitthreading
# threads_per_chain <- 1

#set threads for multithreading
Sys.setenv(STAN_NUM_THREADS = threads_per_chain)


time_cmdstan <- system.time({
  m_col1_prior_cmdstan <- brm(
    dplagoal ~ plaavspeed_c +
    factor(player) +      
    factor(gamenu) +     
    factor(section),
    data    = data_col1,
    family  = bernoulli(link = "logit"),
    #sample_prior = "only",
    prior   = priors_fixef_col1,
    iter    = 1000, warmup = 500,
    chains  = brms_chains, cores = brms_cores,
    backend = "cmdstanr",
    threads = threading(threads = threads_per_chain),
    refresh = 0
  )
})

time_cmdstan
```


```{r}
rbind(
  rstan    = time_rstan,
  cmdstanr = time_cmdstan
)
```


# READ ME Lection:
There are two components that determine runtime. The first is compilation, which translates the model specification(formula and distribution types) into machine-executable code. The second is sampling, during which the MCMC algorithm explores the posterior distribution. 
Once a model is compiled, changes to the data or prior hyperparameters do not require recompilation and only affect sampling time. 
CmdStan typically incurs higher one-time compilation costs but achieves faster sampling. In contrast, rstan usually compiles more quickly but has slower sampling performance. 
-> For models with many predictors or complex structures, sampling time dominates total runtime, making CmdStan the preferred backend.

Within-chain threading introduces overhead and is only beneficial once the per-iteration computational cost of the model is sufficiently large, which is typically the case for complex or high-dimensional models.
-> For models with many predictors or complex structures, multithreading makes sense.

Idea: compile once with tiny model and sample then with more iterations and stuff?




# Teste Laufzeit am größten modell

```{r}
data_col4 <- data_base %>%
  drop_na(
    dplagoal, plaavspeed,
    dhomegame, rank_gap,
    player_cumsections, player_cumsections_2,
    scoregap_pre, dteamgoal_pre, dplagoal_pre,
    dteamred_pre, dteaminj_sec_pre,
    team_avespeed_pre,
    team_crowd, opp_crowd,
    p_odds_team, p_odds_draw
  ) %>%
  mutate(
    plaavspeed_c           = plaavspeed - mean(plaavspeed, na.rm = TRUE),
    rank_gap_c             = rank_gap - mean(rank_gap, na.rm = TRUE),
    player_cumsections_c   = player_cumsections - mean(player_cumsections, na.rm = TRUE),
    player_cumsections_2_c = player_cumsections_2 - mean(player_cumsections_2, na.rm = TRUE),
    scoregap_pre_c         = scoregap_pre - mean(scoregap_pre, na.rm = TRUE),
    team_avespeed_pre_c    = team_avespeed_pre - mean(team_avespeed_pre, na.rm = TRUE),
    team_crowd_c           = team_crowd - mean(team_crowd, na.rm = TRUE),
    opp_crowd_c            = opp_crowd - mean(opp_crowd, na.rm = TRUE),
    p_odds_team_c          = p_odds_team - mean(p_odds_team, na.rm = TRUE),
    p_odds_draw_c          = p_odds_draw - mean(p_odds_draw, na.rm = TRUE)
  )


### Column 4: plus crowd & betting odds
### dplagoal ~ plaavspeed_c + ... + team_crowd_c + opp_crowd_c + p_odds_team_c + p_odds_draw_c + factors
priors_fixef_col3 <- c(
  priors_fixef_base,
  set_prior("normal(0, 0.4)", class = "b", coef = "plaavspeed_c"),
  set_prior("normal(0, 0.4)", class = "b", coef = "dhomegame"),
  set_prior("normal(0, 0.4)", class = "b", coef = "rank_gap_c"),
  set_prior("normal(0, 0.4)", class = "b", coef = "player_cumsections_c"),
  set_prior("normal(0, 0.4)", class = "b", coef = "player_cumsections_2_c"),
  set_prior("normal(0, 0.4)", class = "b", coef = "scoregap_pre_c"),
  set_prior("normal(0, 0.4)", class = "b", coef = "dteamgoal_pre"),
  set_prior("normal(0, 0.4)", class = "b", coef = "dplagoal_pre"),
  set_prior("normal(0, 0.4)", class = "b", coef = "dteamred_pre"),
  set_prior("normal(0, 0.4)", class = "b", coef = "dteaminj_sec_pre"),
  set_prior("normal(0, 0.4)", class = "b", coef = "team_avespeed_pre_c")
)

priors_fixef_col4 <- c(
  priors_fixef_col3,  # all priors from col3 +
  set_prior("normal(0, 0.4)", class = "b", coef = "team_crowd_c"),
  set_prior("normal(0, 0.4)", class = "b", coef = "opp_crowd_c"),
  set_prior("normal(0, 0.4)", class = "b", coef = "p_odds_team_c"),
  set_prior("normal(0, 0.4)", class = "b", coef = "p_odds_draw_c")
)
```


## Standard

```{r}
time_rstan <- system.time({
  m_col4 <- brm(
    dplagoal ~
      plaavspeed_c +
      dhomegame + rank_gap_c +
      player_cumsections_c + player_cumsections_2_c +
      scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
      dteamred_pre + dteaminj_sec_pre +
      team_avespeed_pre_c +
      team_crowd_c + opp_crowd_c +
      p_odds_team_c + p_odds_draw_c +
      factor(player) +      
      factor(gamenu) +     
      factor(section), 
    data    = data_col4,
    family  = bernoulli(link = "logit"),
    prior   = priors_fixef_col4,
    chains  = brms_chains, cores = brms_cores,
    iter = 2000, warmup = 1000,
    backend = "rstan",
    save_pars = save_pars(all = TRUE)   # ensures log-lik etc. are stored
  )
})

time_rstan
```

## CMD Backend

```{r}
time_cmdstan <- system.time({
  m_col4 <- brm(
    dplagoal ~
      plaavspeed_c +
      dhomegame + rank_gap_c +
      player_cumsections_c + player_cumsections_2_c +
      scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
      dteamred_pre + dteaminj_sec_pre +
      team_avespeed_pre_c +
      team_crowd_c + opp_crowd_c +
      p_odds_team_c + p_odds_draw_c +
      factor(player) +      
      factor(gamenu) +     
      factor(section), 
    data    = data_col4,
    family  = bernoulli(link = "logit"),
    prior   = priors_fixef_col4,
    chains  = brms_chains, cores = brms_cores,
    iter = 2000, warmup = 1000,
    backend = "cmdstanr",
    threads = threading(threads = threads_per_chain),
    save_pars = save_pars(all = TRUE)   # ensures log-lik etc. are stored
  )
})

time_cmdstan #43235.63 sekunden ~12h

```

## QR

```{r}
time_QR <- system.time({
  m_col4 <- brm(
    bf(
    dplagoal ~
      plaavspeed_c +
      dhomegame + rank_gap_c +
      player_cumsections_c + player_cumsections_2_c +
      scoregap_pre_c + dteamgoal_pre + dplagoal_pre +
      dteamred_pre + dteaminj_sec_pre +
      team_avespeed_pre_c +
      team_crowd_c + opp_crowd_c +
      p_odds_team_c + p_odds_draw_c +
      factor(player) +      
      factor(gamenu) +     
      factor(section), 
      decomp = "QR"),
    data    = data_col4,
    family  = bernoulli(link = "logit"),
    prior   = priors_fixef_col4,
    chains  = brms_chains, cores = brms_cores,
    iter = 2000, warmup = 1000,
    backend = "cmdstanr",
    threads = threading(threads = threads_per_chain),
    save_pars = save_pars(all = TRUE)   # ensures log-lik etc. are stored
  )
})

time_QR  

```


```{r}
rbind(
  cmdstanr = time_cmdstan,
  QR  = time_QR
)
```

#READ ME Lection: QR ist viel schneller besonders bei factor dummies wegen correlationen. Aber die coefficients und die priors sind nicht mehr brauchbar.


## compare results with OG model

```{r}
# find out how to extract the QR "weights"

library(brms)
library(posterior)

safe <- function(expr) {
  tryCatch(expr, error = function(e) paste("ERROR:", conditionMessage(e)))
}

cat("\n--- BASIC ---\n")
print(safe(class(m_col4)))
print(safe(m_col4$family))

cat("\n--- FORMULA (safe) ---\n")
# formula(m_col4) is usually safe; avoid brmsformula printing methods
print(safe(formula(m_col4)))

cat("\n--- VARIABLES() (first 80) ---\n")
v <- safe(variables(m_col4))
print(v)
if (!is.character(v)) stop("variables(m_col4) failed; cannot proceed.")

cat("\n--- bQ_ PARAMETERS ---\n")
bQ <- grep("^bQ_", v, value = TRUE)
cat("Count bQ_ params:", length(bQ), "\n")
print(head(bQ, 80))

cat("\n--- STANDATA NAMES (QR/R candidates) ---\n")
sd <- safe(standata(m_col4))
if (!is.list(sd)) stop("standata(m_col4) failed; cannot proceed.")

nms <- names(sd)
cand <- nms[grepl("qr|QR|R_|\\bR\\b|decomp|Q_|Rinv|inv", nms)]
cat("Candidate names count:", length(cand), "\n")
print(cand)

cat("\n--- DIMENSIONS OF CANDIDATES ---\n")
dims <- lapply(sd[cand], function(x) if (is.null(dim(x))) NA else dim(x))
print(dims)

cat("\n--- ALL objects in standata() that have dim() (first 60 names) ---\n")
mat_names <- nms[vapply(sd, function(x) !is.null(dim(x)), logical(1))]
cat("Total dim() objects:", length(mat_names), "\n")
print(head(mat_names, 60))

cat("\n--- DONE ---\n")


```

```{r}
library(brms)
library(posterior)
library(MASS)  # for ginv()

sd <- standata(m_col4)
X  <- sd$X

draws <- as_draws_df(m_col4)

# bQ indices robustly (no "NAs introduced by coercion")
bQ_cols <- grep("^bQ\\[[0-9]+\\]$", names(draws), value = TRUE)
if (!length(bQ_cols)) stop("No bQ[...] parameters found.")

bQ_idx  <- as.integer(sub("^bQ\\[([0-9]+)\\]$", "\\1", bQ_cols))
ord     <- order(bQ_idx)
bQ_cols <- bQ_cols[ord]
K       <- length(bQ_cols)

# Align X to K (Intercept often not in QR block)
if (ncol(X) == K + 1) {
  is_const <- apply(X, 2, function(z) all(abs(z - z[1]) < 1e-12))
  X <- X[, !is_const, drop = FALSE]
  if (ncol(X) != K) X <- X[, 1:K, drop = FALSE]  # last-resort alignment
}
stopifnot(ncol(X) == K)

# Recompute R from X
R <- qr.R(qr(X))

# theta draws
theta <- as.matrix(draws[, bQ_cols, drop = FALSE])  # draws x K

# Back-transform using pseudo-inverse (works even if R is singular)
R_pinv <- MASS::ginv(R)
beta   <- t(R_pinv %*% t(theta))                    # draws x K

# Summarize
beta_summary <- data.frame(
  j        = 1:K,
  Estimate = colMeans(beta),
  Est.Error = apply(beta, 2, sd),
  Q2.5     = apply(beta, 2, quantile, probs = 0.025),
  Q97.5    = apply(beta, 2, quantile, probs = 0.975)
)

print(beta_summary)



```

